```{r echo=FALSE, message=FALSE, warning=FALSE}
# Using average ratio between elections
fe_lm_0 <- feols(vote_change ~
                   emigration_election_year_per_1000 +
                   volatility + gdp + unemployment |
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_1 <- feols(vote_change ~
                   emigration_election_year_per_1000 +
                   average_ratio_schools_election_year +
                   volatility + gdp + unemployment |
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_2 <- feols(vote_change ~
                   emigration_election_year_per_1000 +
                   average_ratio_classrooms_election_year +
                   volatility + gdp + unemployment |
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_3 <- feols(vote_change ~
                   emigration_election_year_per_1000 +
                   average_ratio_hospitals_election_year +
                   volatility + gdp + unemployment |
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_4 <- feols(vote_change ~
                   emigration_election_year_per_1000 +
                   average_ratio_third_places_election_year +
                   volatility + gdp + unemployment |
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_5 <- feols(vote_change ~
                   emigration_election_year_per_1000 +
                   average_ratio_schools_election_year +
                   average_ratio_hospitals_election_year +
                   average_ratio_third_places_election_year +
                   volatility + gdp + unemployment |
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_6 <- feols(vote_change ~
                   emigration_election_year_per_1000 +
                   average_ratio_classrooms_election_year +
                   average_ratio_hospitals_election_year +
                   average_ratio_third_places_election_year +
                   volatility + gdp + unemployment |
                   nuts2016 + year,
                 data = anti_incumbent_vote)
```

```{r results-emigration-incumbent-vote-share, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Emigration interacted with Service Cuts on Incumbent Vote Share; DV: Change in Incumbent Vote Share; Model Specification A"}

# groups_list <- list(
#   "Main Variables" = c("Emigration since last election", "Avg. Children per School", "Avg. Children per Classroom", "Avg. People per Hospital", "Avg. People per Third Place"),
#   "Control Variables" = c("Volatility", "GDP per capita", "Unemployment"))

model_table <- modelsummary(list("Model 0" = fe_lm_0,
                                 "Model 1" = fe_lm_1,
                                 "Model 2" = fe_lm_2,
                                 "Model 3" = fe_lm_3,
                                 "Model 4" = fe_lm_4,
                                 "Model 5" = fe_lm_5,
                                 "Model 6" = fe_lm_6),
             title = "migration interacted with Service Cuts on Incumbent Vote Share; DV: Change in Incumbent Vote Share; Model Specification A",
             coef_map = c(`average_ratio_schools_election_year` = "Avg. Children per School",
                          `average_ratio_classrooms_election_year` = "Avg. Children per Classroom",
                          `average_ratio_hospitals_election_year` = "Avg. People per Hospital",
                          `average_ratio_third_places_election_year` = "Avg. People per Third Place",
                          emigration_election_year_per_1000 = "Emigration since last election",
                          volatility = "Volatility",
                          gdp = "GDP per capita",
                          unemployment = "Unemployment"),
             # groups = groups_list,
             statistic = "p.value",
             gof_omit = "AIC|BIC|RMSE|Std.Errors|FE",
             stars = T,
             notes = "Fixed Effects: NUTS regions and years. Standard errors are grouped by NUTS3 region. p-values shown in parentheses.")

model_table %>%
  kable_styling(font_size = 7)
```

<!-- description of model specification A results -->
Table \@ref(tab:results-emigration-incumbent-vote-share) displays the results of Model Specification A. This models the average change of number of people per institution seperately on incumbent support, while controlling for external emigration, electoral volatility, GDP per capita and unemployment while applying NUTS3 regions and years as fixed effects. Models 1 through 4 contain each service cut seperately, while Models 5 and 6 combine all service cuts together. Model 5 uses the number of schools and Model 6 applies the classroom proxy to operationalise service cuts to schools. Model 0 does not include any service cuts and is included in Table \@ref(tab:results-emigration-incumbent-vote-share) as a reference model.

The service cuts by themselves (Models 1 through 4) show mixed results. Service cuts to schools (school buildings or classrooms) display negative coefficients, meaning that an increase of children per school or classroom is associated with a decrease in incumbent support. While the coefficient for Model 2 is not statistically significant, the average number of children per school (Model 1) displays borderline statistical significance ($p = 0.103$). We can interpret this coefficient to mean that while controlling for emigration, electoral volatility, GDP per capita and unemployment, for a given Romanian NUTS3 region, in a year where the average rate of children per school is one child higher that it typically is for that region, we would expect the incumbent vote share to decrease by $0.093$ percentage points more than is typical for that region. The coefficient value can also be interpreted the following way. If a school grows by eleven children, then this is associated with an incumbent vote share decrease of around $1$ percentage point, while controlling for the above mentioned variables.

The change in number of hospitals has no effect, and the change in number of third places shows a positive effect. This indicates that an increase in the number of third places is associated with an increase in incumbent vote share. The results in Models 3 and 4 are not statistically significant at conventional levels.

<!-- one vs all independent variables -->
By combining the different service cuts in Models 5 and 6, we can examine what the unique contributions are of each service cut on incumbent vote change. However, the results from Models 5 and 6 are very similar to the results from Models 1 through 4. Again, only school service cuts show negative associations, but these results are not statistically significant and their coefficient strengths change marginally. However, this stability can be interpreted positively. These unchanged coefficients indicate that the effect of each service cut on incumbent vote share change is relatively independent of each other. In other words, the different service cuts capture distinct aspects of the variance of incumbent vote share change without overlapping with each other. Thus I interpret the coefficient stability to show that the model is robust.

<!-- control variable coefficient results: do control variables align with theoretical expectations? -->
I move on to report on the control variable coefficient results. I acknowledge that control variables are not in the primary interest, however they can contribute to evidence of effects on incumbent vote support [@hunermundNuisanceControlVariables2023 4]. I begin by discussing the coefficient of the control variable emigration and then its varying coefficient strength in the different models. Across all models, we can observe that an increase in emigration is associated with a decrease in incumbent vote share. This fulfils our theoretical expectations and fits into the causal mechanism of emigration causing grievances, which in turn decreases incumbent support.

<!-- emigration change in coefficient strength over models: omitted variable bias -->
Depending on which service cut is applied, the emigration coefficient varies in strength. This hints at omitted variable bias in Model 0, the reference model. We observe the biggest change in strength when school service cuts are added, using the children per school operationalisation (Model 1). Here the emigration coefficient increases by $0.339$, from $-5.5$ to $-5.161$. This indicates that Model 0 was underspecified and suffered from omittied variables bias, overstating the effect of emigration on incumbent support. This is then corrected when school service cuts are added to the model. Same goes for Models 2, 4 and 6, as these models also see a decrease in coefficient size, however in smaller amounts compared to Model 1.

Two of the three further control variables display coefficients that align with the theoretical assumptions and are statistically significant across all models. Both economic variables, GDP per capital and unemployment, show coefficients that align with theoretical expectations. An increase in per capita GDP is associated with an increase in incumbent support, which can be interpreted that incumbent parties are rewarded for a good state of the economy. On the other hand, we see that an increase in unemployment is associated with a decrease in incumbent support. This suggests that voters punish incumbent parties for poor economic performance. Electoral volatility shows a positive coefficient throughout all models, indicating an increase in volatility is associated with an increase in incumbent support. This does not make intuitive sense, as we would expect high levels of volatility to be associated with a decrease in incumbent support, as incumbent parties would be voted out of office and replaced with new parties.

<!-- R2 values; https://stats.stackexchange.com/a/401331 -->
One difference between the service cuts-combining models (Models 5 and 6) and the other models (Models 1 through 4) are the increased Adjusted Within R-Squared values. We can interpret this value as the variation in incumbent vote share change within Romanian NUTS3 regions over time that is captured by the models. E.g., Model 5 displays the highest Adjusted Within R-Squared, with $28.6\%$ of the variation being explained by the model. This indicates that combining all service cuts as independent variables makes the models fit the data better than using the service cuts by themselves. This increase in within groups explanatory power hints at the possibility that all these different service cuts contribute to influencing incumbent support.
