```{r echo=FALSE, message=FALSE, warning=FALSE}
# With control variables
# Using average ratio between elections
fe_lm_0 <- feols(vote_change ~
                   emigration_election_year_per_1000 +
                   volatility + gdp + unemployment |
                   nuts2016 + year,
                 data = anti_incumbent_vote)
#this is the one
fe_lm_1 <- feols(vote_change ~
                   average_ratio_schools_election_year *
                   emigration_election_year_per_1000 +
                   volatility + gdp + unemployment |
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_2 <- feols(vote_change ~
                   average_ratio_classrooms_election_year *
                   emigration_election_year_per_1000 +
                   volatility + gdp + unemployment |
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_3 <- feols(vote_change ~
                   average_ratio_hospitals_election_year *
                   emigration_election_year_per_1000 +
                   volatility + gdp + unemployment |
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_4 <- feols(vote_change ~
                   average_ratio_third_places_election_year *
                   emigration_election_year_per_1000 +
                   volatility + gdp + unemployment |
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_5 <- feols(vote_change ~
                   average_ratio_schools_election_year *
                   emigration_election_year_per_1000 +
                   
                   average_ratio_hospitals_election_year *
                   emigration_election_year_per_1000 +

                   average_ratio_third_places_election_year *
                   emigration_election_year_per_1000 +
                   
                   volatility + gdp + unemployment |
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_6 <- feols(vote_change ~
                   average_ratio_classrooms_election_year *
                   emigration_election_year_per_1000 +
                   
                   average_ratio_hospitals_election_year *
                   emigration_election_year_per_1000 +
                   
                   average_ratio_third_places_election_year *
                   emigration_election_year_per_1000 +
                   
                   volatility + gdp + unemployment |
                   nuts2016 + year,
                 data = anti_incumbent_vote)
```

```{r results-emigration-incumbent-vote-share-control-interaction, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Emigration interacted with Service Cuts on Incumbent Vote Share with Control Variables; DV: Change in Incumbent Vote Share; Model Specification B"}
# new_names <- c(
#   "average_ratio_classrooms_election_year:emigration_election_year_per_1000" = "Classrooms-Emigration Interaction",
#   "emigration_election_year_per_1000:average_ratio_hospitals_election_year" = "Hospitals-Emigration Interaction",
#   "emigration_election_year_per_1000:average_ratio_third_places_election_year" = "Third Places-Emigration Interaction",
#   "volatility" = "Economic Volatility",
#   "gdp" = "Gross Domestic Product",
#   "unemployment" = "Unemployment Rate"
# )
# 
# modelsummary(fe_lm_6, coef_map = new_names)

model_table <- modelsummary(list("Model 0" = fe_lm_0,
                                 "Model 1" = fe_lm_1,
                                 "Model 2" = fe_lm_2,
                                 "Model 3" = fe_lm_3,
                                 "Model 4" = fe_lm_4
                                 #"Model 5" = fe_lm_5,
                                 #"Model 6" = fe_lm_6
                                 ),
              title = "Emigration interacted with Service Cuts on Incumbent Vote Share with Control Variables; DV: Change in Incumbent Vote Share; Model Specification B",
             coef_map = c("average_ratio_schools_election_year:emigration_election_year_per_1000" = "Avg. Children per School × Emigration since last election",
                          "average_ratio_classrooms_election_year:emigration_election_year_per_1000" = "Avg. Children per Classroom × Emigration since last election",
                          "emigration_election_year_per_1000:average_ratio_hospitals_election_year" = "Avg. People per Hospital × Emigration since last election",
                          "average_ratio_hospitals_election_year:emigration_election_year_per_1000" = "Avg. People per Hospital × Emigration since last election",
                          "emigration_election_year_per_1000:average_ratio_third_places_election_year" = "Avg. People per Third Place × Emigration since last election",
                          "average_ratio_third_places_election_year:emigration_election_year_per_1000" = "Avg. People per Third Place × Emigration since last election",
                          "emigration_election_year_per_1000" = "Emigration since last election",
                          "volatility" = "Volatility",
                          "gdp" = "GDP per capita",
                          "unemployment" = "Unemployment"),
             gof_omit = "AIC|BIC|RMSE|Std.Errors|FE",
             statistic = "p.value",
             stars = T,
             notes = "Fixed Effects: NUTS regions and years. Standard errors are grouped by NUTS3 region. p-values shown in parentheses.")

model_table %>%
  kable_styling(font_size = 7)
```

<!-- description of model specification B results -->
Table \@ref(tab:results-emigration-incumbent-vote-share-control-interaction) displays the results of Model Specification B, which models the interaction between average change of number of people per institution on incumbent support, while controlling for external emigration, electoral volatility, GDP per capita and unemployment and applying NUTS3 regions and year fixed effects. The results from these models tell us how a service cut and the change in incumbent vote share are associated, given a specific emigration rate. The models in Table \@ref(tab:results-emigration-incumbent-vote-share-control-interaction) also include coefficients for service cuts on their own, in order to measure the direct effect of service cuts on incumbent vote share change. However, these coefficients have been hidden in Table \@ref(tab:results-emigration-incumbent-vote-share-control-interaction) in order to focus on the interaction terms. Furthermore, the interpretation of the service cut coefficients by themselves are not meaningful. Table \@ref(tab:all-results-emigration-incumbent-vote-share-control-interaction) in the appendix includes all variable coefficients. Model 0 does not include any service cuts and is included in Table \@ref(tab:results-emigration-incumbent-vote-share-control-interaction) as a reference model.

Similar to the results shown in Table \@ref(tab:results-emigration-incumbent-vote-share), service cuts display mixed results when interacting with emigration. When modelling service cuts conditioned on emigration separately on incumbent vote share change, only school service cuts (Model 1, operationalised using school buildings) displays a coefficient that aligns with the theoretical expectations, i.e., a negative coefficient, and is borderline statistically significant ($p = 0.112$). This result shows that the effect of service cuts on incumbent vote share change changes when the external emigration rate differs. The only other coefficient result to meet the theoretical expectations are the hospital service cuts when interacted with emigration (Model 3). However, this coefficient is not statistically significant at conventional levels ($p = 0.200$) and the association is small in magnitude ($-0.001$).

<!-- Interpreting interaction terms https://theeffectbook.net/ch-StatisticalAdjustment.html#interaction-terms -->
I continue by interpreting Model 1’s negative interaction coefficient. The base effect of school service cuts on incumbent vote share change, without taking emigration into account, equals $0.045$.^[See Table \@ref(tab:all-results-emigration-incumbent-vote-share-control-interaction) in the Appendix]. This means that in a year where the average rate of children per school is one child higher that it typically is for that region, incumbent vote share change *increases* by $0.045$ percentage points more than is typical for that region, holding all other variables constant. The interaction term modifies this effect based on the emigration rate. For each one-unit increase in emigration, the effect of school service cuts on incumbent vote share change *decreases* by $0.058$ percentage points. In order to calculate the total effect of a one child per school increase when emigration also increases by one person per 1000, I sum the base effect of service cuts and the interaction effect: $0.045 + -0.058 = -0.013$. This means that when emigration is added as an interaction term, the effect of school service cuts on incumbent vote share change shifts from a positive effect of $0.045$ percentage points to a negative effect of $-0.013$ percentage points. In short, in Romanian NUTS3 regions where the emigration rate increases by one unit, an increase in children per school is associated with a *decrease* in the incumbent vote share change.

When we look at the interaction term between the school service cuts operationalised with children per classroom together with emigration (Model 2), we see, despite Model 2’s positive coefficient in Table \@ref(tab:results-emigration-incumbent-vote-share-control-interaction), a similar result. The base effect of average children per classroom is $-4.753$.^[See Table \@ref(tab:all-results-emigration-incumbent-vote-share-control-interaction) in the Appendix] The interaction term equals $1.124$. I calculate the sum of the base effect and interaction term: $-4.753 + 1.124 = -3.629$. Thus when emigration is added as an interaction term, the effect of school service cuts increases slightly but still stays below zero, indicating a negative effect on incumbent support. However, this interaction is not statistically significant, thus these results should be interpreted with caution.

In sum, the negative effect of school service cuts on the change in incumbent vote share becomes more pronounced as emigration increases. In other words, in Romanian NUTS3 regions with higher emigration rates, the impact of school closures on reducing incumbent vote share is stronger. However, due to the borderline statistical significance, the results must be interpreted with caution.

Models 2, 3 and 4 do not display statistically significant interaction effects. Furthermore, the effects are either very small in magnitude (Model 3) or display a positive coefficient and thus do not align with the theoretical expectations (Models 2 and 4).

<!-- control variables -->
As in Table \@ref(tab:results-emigration-incumbent-vote-share), we can observe that higher emigration rates are associated with a decrease in incumbent vote share. This aligns with the theoretical expectations.

<!-- emigration change in coefficient strength over models: omitted variable bias -->
When comparing coefficient strengths of the emigration control variable, we see that Model 1 displays a large drop in strength compared to the reference Model 0. The coefficient increases by $2.271$, from $-5.5$ to $-3.229$. As in Model 1 in Table \@ref(tab:results-emigration-incumbent-vote-share), this again indicates that Model 0 in Table \@ref(tab:results-emigration-incumbent-vote-share-control-interaction) suffers from omitted variable bias and that adding the interaction between school service cuts and emigration corrects this. A similar coefficient increase can be observed in Models 2 and 4, however to a lesser extent than Model 1.

<!-- control variable coefficient results: do control variables align with theoretical expectations? -->
When observing the other control variables in all models in Table \@ref(tab:results-emigration-incumbent-vote-share-control-interaction), we see similar coefficients as in Table \@ref(tab:results-emigration-incumbent-vote-share). Two of the three further control variables display coefficients that align with the theoretical assumptions and are statistically significant across all models. The one control variable that does not align with the theoretical expectation is again electoral volatility, as it shows positive coefficients across all models. This indicates that an increase in volatility is associated with an increase in incumbent vote share change, not an intuitive result.

Because electoral volatility displays unexpected results as a control variable across all models and in both model specifications, this may point towards an error in its operationalisation or in the underlying data. \label{passage:volatility-issue} This issue is discussed in the Conclusion.

<!-- R2 values; https://stats.stackexchange.com/a/401331 -->
As with the models in Table \@ref(tab:results-emigration-incumbent-vote-share), we can observe some variation in the Adjusted Within R-Squared value across the different models. Their values are roughly at the same levels as the models in Table \@ref(tab:results-emigration-incumbent-vote-share), indicating that both the models with and without interaction terms capture a similar amount of variation in incumbent vote share change within Romanian NUTS3 regions over time.
