
```{r echo=FALSE, message=FALSE, warning=FALSE}
# With control variables
# Using average ratio between elections
fe_lm_0 <- feols(vote_change ~
                   emigration_election_year_per_1000 +
                   volatility + gdp + unemployment |
                   nuts2016 + year,
                 data = anti_incumbent_vote)
#this is the one
fe_lm_1 <- feols(vote_change ~
                   average_ratio_schools_election_year *
                   emigration_election_year_per_1000 +
                   volatility + gdp + unemployment |
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_2 <- feols(vote_change ~
                   average_ratio_classrooms_election_year *
                   emigration_election_year_per_1000 +
                   volatility + gdp + unemployment |
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_3 <- feols(vote_change ~
                   average_ratio_hospitals_election_year *
                   emigration_election_year_per_1000 +
                   volatility + gdp + unemployment |
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_4 <- feols(vote_change ~
                   average_ratio_third_places_election_year *
                   emigration_election_year_per_1000 +
                   volatility + gdp + unemployment |
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_5 <- feols(vote_change ~
                   average_ratio_schools_election_year *
                   emigration_election_year_per_1000 +
                   
                   average_ratio_hospitals_election_year *
                   emigration_election_year_per_1000 +

                   average_ratio_third_places_election_year *
                   emigration_election_year_per_1000 +
                   
                   volatility + gdp + unemployment |
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_6 <- feols(vote_change ~
                   average_ratio_classrooms_election_year *
                   emigration_election_year_per_1000 +
                   
                   average_ratio_hospitals_election_year *
                   emigration_election_year_per_1000 +
                   
                   average_ratio_third_places_election_year *
                   emigration_election_year_per_1000 +
                   
                   volatility + gdp + unemployment |
                   nuts2016 + year,
                 data = anti_incumbent_vote)
```

```{r results-emigration-incumbent-vote-share-control-interaction, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Emigration interacted with Service Cuts on Incumbent Vote Share with Control Variables; DV: Change in Incumbent Vote Share; Model Specification B"}
# new_names <- c(
#   "average_ratio_classrooms_election_year:emigration_election_year_per_1000" = "Classrooms-Emigration Interaction",
#   "emigration_election_year_per_1000:average_ratio_hospitals_election_year" = "Hospitals-Emigration Interaction",
#   "emigration_election_year_per_1000:average_ratio_third_places_election_year" = "Third Places-Emigration Interaction",
#   "volatility" = "Economic Volatility",
#   "gdp" = "Gross Domestic Product",
#   "unemployment" = "Unemployment Rate"
# )
# 
# modelsummary(fe_lm_6, coef_map = new_names)

model_table <- modelsummary(list("Model 0" = fe_lm_0,
                                 "Model 1" = fe_lm_1,
                                 "Model 2" = fe_lm_2,
                                 "Model 3" = fe_lm_3,
                                 "Model 4" = fe_lm_4
                                 #"Model 5" = fe_lm_5,
                                 #"Model 6" = fe_lm_6
                                 ),
              title = "Emigration interacted with Service Cuts on Incumbent Vote Share with Control Variables; DV: Change in Incumbent Vote Share; Model Specification B",
             coef_map = c("average_ratio_schools_election_year:emigration_election_year_per_1000" = "Avg. Children per School × Emigration since last election",
                          "average_ratio_classrooms_election_year:emigration_election_year_per_1000" = "Avg. Children per Classroom × Emigration since last election",
                          "emigration_election_year_per_1000:average_ratio_hospitals_election_year" = "Avg. People per Hospital × Emigration since last election",
                          "average_ratio_hospitals_election_year:emigration_election_year_per_1000" = "Avg. People per Hospital × Emigration since last election",
                          "emigration_election_year_per_1000:average_ratio_third_places_election_year" = "Avg. People per Third Place × Emigration since last election",
                          "average_ratio_third_places_election_year:emigration_election_year_per_1000" = "Avg. People per Third Place × Emigration since last election",
                          "emigration_election_year_per_1000" = "Emigration since last election",
                          "volatility" = "Volatility",
                          "gdp" = "GDP per capita",
                          "unemployment" = "Unemployment"),
             gof_omit = "AIC|BIC|RMSE|Std.Errors|FE",
             statistic = "p.value",
             stars = T,
             notes = "Fixed Effects: NUTS regions and years. Standard errors are grouped by NUTS3 region. p-values shown in parentheses.")

model_table %>%
  kable_styling(font_size = 7)
```

<!-- description of model specification B results -->
- Table \@ref(tab:results-emigration-incumbent-vote-share-control-interaction) displays the results of modelling the interaction between average change of number of people per institution on incumbent support, while controlling for emigration, volatility, GDP und unemployment and applying NUTS3 region and year fixed effects; This is the implementation of Model Specification B.
- The results from these models tell us what the effect of a service cut is, given a specific emigration rate
- The models in Table \@ref(tab:results-emigration-incumbent-vote-share-control-interaction) also include coefficients for service cuts on their own, in order to measure the direct effect of service cuts on incumbent vote share change. However, they have been removed from the table to keep it tidy and because the focus is on the interaction terms and the interpretation of the service cut coefficients by themselves are not meaningful.
- Model 0 does not include any service cuts variables and is included in Table \@ref(tab:results-emigration-incumbent-vote-share-control-interaction) as a reference model

- As in Table \@ref(tab:results-emigration-incumbent-vote-share), the service cuts when interacted with emigration display mixed results. 
- When modelling service cuts conditioned on emigration seperately on incumbent vote share change, only service cuts to schools (Model 1 operationalised using school buildings) displays a coefficient that aligns with the theoretical expectations, i.e., a negative coefficient, and is (borderline) statistically significant at $p = 0.112$.
- This statistically significant term hints that the effect of service cuts on incumbent vote share change differs between different emigration rates
- The only other coefficient to meet the theoretical expectations is that of hospital service cuts when interacted with emigration (Model 3). However, this coefficient is not statistically significant at conventional levels ($p = 0.200$) and the coefficient’s size at $-0.001$ is very small.

<!-- Interpreting interaction terms https://theeffectbook.net/ch-StatisticalAdjustment.html#interaction-terms -->
- What does Model 1’s negative interaction coefficient mean? We take a closer look
- The base effect of school service cuts on incumbent vote share change, without taking the emigration into account, equals $0.045$ (not shown in Table \@ref(tab:results-emigration-incumbent-vote-share-control-interaction)). This means in a year where the average rate of children per school is one child higher that it typically is for that region, incumbent vote share change increases by $0.045$ more than is typical for that region, holding all other variables constant
- The interaction term modifies this effect based on the emigration rate. For each one-unit increase in emigration, the effect of school service cuts on incumbent vote share change decreases by $0.058$
- In order to calculate the total effect of a one child per school increase when emigration also increases by one person per 1000, I add the base effect of service cuts and the interaction effect together: $0.045 - 0.058 = -0.013$.
- This means that when emigration is added as an interaction term, the effect of school service cuts on incumbent vote share change shifts from a positive effect of $0.045$ to a negative effect of $0.013$.
- In other words, in NUTS3 regions where the emigration rate is higher by one unit, an increase in children per school is associated with a decrease in incumbent vote share change, rather than an increase.
- In sum, “the negative effect of service cuts on the change in incumbent vote share becomes more pronounced as the emigration rate increases. In other words, in Romanian NUTS3 regions with higher emigration rates, the impact of less schools on reducing incumbent vote share is stronger."
- Models 2, 3 and 4 do not display statistially significant effects. Also, their effects are very small or do not align with my theoretical expectations

<!-- emigration change in coefficient strength over models: omitted variable bias -->
- Model 1 displays a large drop in coefficient strength of the control variable emigration
- Compared to the reference Model 0, the coefficient increases by $2.271$, from $-5.5$ to $-3.229$. 
- As in Model 1 in Table \@ref(tab:results-emigration-incumbent-vote-share), this again indicates that Model 0 in Table \@ref(tab:results-emigration-incumbent-vote-share-control-interaction) suffered from omitted variable bias and that adding the interaction between school service cuts and emigration corrects this bias
- A similar coefficient increase can be observed in Models 2 and 4, however in smaller amounts compared to Model 1.

<!-- R2 values; https://stats.stackexchange.com/a/401331 -->
- As with the models in Table \@ref(tab:results-emigration-incumbent-vote-share), we can observe some variation in the Adjust Within R-Squared value across the different models. 
- Their values are roughly at the same levels as the models in Table \@ref(tab:results-emigration-incumbent-vote-share), indicating that both the models with and without interaction terms capture roughly the same variation in incumbent vote share change within Romanian NUTS3 regions over time.

<!-- control variable coefficient results: do control variables align with theoretical expectations? -->
- When looking at the control variables in the models in Table \@ref(tab:results-emigration-incumbent-vote-share-control-interaction), we see similar coefficients as in Table \@ref(tab:results-emigration-incumbent-vote-share).
- Again, three of the four control variables display coefficients that align with the theoretical assumptions and are statistically significant across all models. 
- The one control variable that does not align with the theoretical expectations is again the Volatility coefficient, as is shows a positive value across all models. This indicates that an increase in volatility is associated with an increase in incumbent vote share change, an assumption that does not make sense intuitively.
