## Results
- Do emigration induced service cuts effect incumbent vote share in Poland and Romania?
- Hypotheses acceptance or rejection
- Discussion of Polish and Romanian service cut models and their respective differences
	- Stronger effect of emigration on incumbent vote share in Romania
	- Generally stronger effects of service cuts in Romania than Poland


```{r echo=FALSE, message=FALSE, warning=FALSE}
# Romania
anti_incumbent_vote <- ned_v_dem_cee %>% 
  dplyr::filter(country == "Romania") %>%
  dplyr::filter(prev_incumbent == T) %>%
  left_join(dplyr::select(ro, year, NUTS_ID, emigration_election_year_per_1000), by = c("year" = "year", "nuts2016" = "NUTS_ID")) %>%
  left_join(dplyr::select(ro_schools, -population), by = c("year" = "year", "nuts2016" = "NUTS_ID")) %>%
  left_join(dplyr::select(ro_hospitals, -population), by = c("year" = "year", "nuts2016" = "NUTS_ID")) %>% 
  left_join(dplyr::select(ro_third_places, -population), by = c("year" = "year", "nuts2016" = "NUTS_ID")) %>% 
  left_join(ro_volatility, by = c("year", "nuts2016")) %>% 
  left_join(ro_gdp, by = c("year" = "year", "nuts2016" = "NUTS_ID"))
```

### Ratios
**Just using Service Cuts (no control variables)**

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Using average ratio between elections
fe_lm_0 <- feols(vote_change ~
                   emigration_election_year_per_1000 |
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_1 <- feols(vote_change ~
                   emigration_election_year_per_1000 *
                   average_ratio_schools_election_year |
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_2 <- feols(vote_change ~
                   emigration_election_year_per_1000 *
                   average_ratio_classrooms_election_year |
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_3 <- feols(vote_change ~
                   emigration_election_year_per_1000 *
                   average_ratio_hospitals_election_year |
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_4 <- feols(vote_change ~
                   emigration_election_year_per_1000 *
                   average_ratio_third_places_election_year |
                   nuts2016 + year,
                 data = anti_incumbent_vote)
```

**With control variables**
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Using average ratio between elections
fe_lm_0 <- feols(vote_change ~
                   emigration_election_year_per_1000 +
                   volatility + gdp |
                   nuts2016 + year,
                 data = anti_incumbent_vote)
#this is the one
fe_lm_1 <- feols(vote_change ~
                   emigration_election_year_per_1000 *
                   average_ratio_schools_election_year +
                   volatility + gdp |
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_2 <- feols(vote_change ~
                   emigration_election_year_per_1000 *
                   average_ratio_classrooms_election_year +
                   volatility + gdp |
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_3 <- feols(vote_change ~
                   emigration_election_year_per_1000 *
                   average_ratio_hospitals_election_year +
                   volatility + gdp |
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_4 <- feols(vote_change ~
                   emigration_election_year_per_1000 *
                   average_ratio_third_places_election_year +
                   volatility + gdp |
                   nuts2016 + year,
                 data = anti_incumbent_vote)
```

