#### Does emigration affect incumbent vote share?

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Poland
anti_incumbent_vote <- ned_v_dem_cee %>% 
  dplyr::filter(country == "Poland") %>%
  dplyr::filter(prev_incumbent == T) %>%
  left_join(dplyr::select(pl, year, NUTS_ID, emigration_election_year_per_1000), by = c("year" = "year", "nuts2016" = "NUTS_ID")) %>%
  left_join(dplyr::select(pl_schools, -population), by = c("year" = "year", "nuts2016" = "NUTS_ID")) %>%
  left_join(dplyr::select(pl_hospitals, -population), by = c("year" = "year", "nuts2016" = "NUTS_ID")) %>% 
  left_join(dplyr::select(pl_third_places, -population), by = c("year" = "year", "nuts2016" = "NUTS_ID"))

fe_lm_pl <- feols(vote_change ~ emigration_election_year_per_1000 | nuts2016 + year, data = anti_incumbent_vote)

# Romania
anti_incumbent_vote <- ned_v_dem_cee %>% 
  dplyr::filter(country == "Romania") %>%
  dplyr::filter(prev_incumbent == T) %>%
  left_join(dplyr::select(ro, year, NUTS_ID, emigration_election_year_per_1000), by = c("year" = "year", "nuts2016" = "NUTS_ID")) %>%
  left_join(dplyr::select(ro_schools, -population), by = c("year" = "year", "nuts2016" = "NUTS_ID")) %>%
  left_join(dplyr::select(ro_hospitals, -population), by = c("year" = "year", "nuts2016" = "NUTS_ID")) %>% 
  left_join(dplyr::select(ro_third_places, -population), by = c("year" = "year", "nuts2016" = "NUTS_ID"))

fe_lm_ro <- feols(vote_change ~ emigration_election_year_per_1000 | nuts2016 + year, data = anti_incumbent_vote)
```

```{r emigration-incumbent-vote-share, echo=FALSE, message=FALSE, warning=FALSE}
model_table <- modelsummary(list("Poland" = fe_lm_pl,
                                 "Romania" = fe_lm_ro),
             title = "DV: Change in Incumbent Vote Share",
             coef_map = c(emigration_election_year_per_1000 = "Emigration since last election"),
             gof_omit = "AIC|BIC|RMSE|Std.Errors|FE",
             stars = T,
             statistic = NULL,
             notes = "Fixed Effects: NUTS regions and years. Standard errors are grouped by NUTS region.")

model_table %>%
  kable_styling(font_size = 10)
```

I examine if there is an association between emigration rates and anti-incumbent voting. Following the theory, emigration cause grievances, which in turn can manifest as a vote against the incumbent party. Thus I expect that higher emigration to be associated with a decrease in incumbent vote share.

As the dependent variable I use the incumbent vote change between national parliamentary elections. The model uses the total emigration rate from a region between two elections as the independent variable. The precise operationalisation of these variables are described in XXX (TO DO). I model Poland and Romania seperately, using a fixed effects OLS estimation. Results for both models are displayed in Table \@ref(tab:emigration-incumbent-vote-share).

Emigration does not have a significant effect on incumbent voting in Poland. Also, the coefficient is positive, which indicates that an increase in emigration is associated with an increase in incumbent vote share. Romania displays a result that aligns better with my theory. Here, higher emigration rates between elections are associated with a relatively strong decrease in incumbent vote share. An increase of one emigration per thousand inhabitants of a region is associated with a decrease of incumbent vote share of about $6.5\%$. The coefficient displays statistical significance at conventional levels. Furthermore, the Romanian model displays relatively high $R^2$ levels, indicating that the model explains a large part of the dependent’s variable overall variance.