### Operationalisation

#### Dependent Variable: Incumbent Vote Share Change
\label{passage:incumbent-vote-share-change} In my analysis, the dependent variable is the change in incumbent vote share. To operationalise this, I pick two consecutive national parliamentary elections, referred to as $E_0$ and $E_1$. Following the data provided by @staffani.lindbergVPartyDatasetV22022 and manual adjustments (see p. \pageref{passage:incumbent-party}), I label each party as either incumbent or non-incumbent at $E_0$. Next, I calculate the change in the party's vote share between $E_0$ and $E_1$. A positive value indicates an increase in support, a negative value signifies a loss in vote share and thus less support.

#### Independent Variable: Service Cuts
I operationalise service cuts based on the assumption that a person's experience of a service is influenced by the number of people per institution, e.g., per school or hospital. When one institution closes down, more people use the remaining institutions, whereby an individual’s experience of the services provided by the remaining institutions deteriorates. This is due to factors such as diminished individual attention, overstretched personnel and longer waiting times.

In short, a service cut, such as the closure of a hospital, results in an increased number of people who use the remaining institutions. This leads to a perceived decrease in service quality. By calculating the ratio of people per institution in a region, I aim to capture the perceived quality of a service. This offers a more nuanced understanding than merely counting the number of institutions in a region. Also, by using this ratio I can account for change in population size, e.g., population decline in regions affected by emigration. Measuring service cuts this way aims to capture the deterioration of service quality and its impact on the development of grievances, which finally lead to a change of voting behaviour.

Service cuts are operationalised by examining the average change of people per institution between two elections. For service cuts to schools, this only includes the primary school aged population in a NUTS region. For service cuts to hospitals and third places, this covers a region’s entire population, as all a region’s inhabitants are potential users of hospitals and third places as services.

Service cut operationalisation can be best understood with an example, as displayed in Table \@ref(tab:ratio-example). The number of primary-aged children per school in a specific NUTS region in a given year equals the ratio. This ratio varies annually and is recorded in the “Yearly Change in Children per School” column. To calculate the average change between two elections, I average the yearly change values starting from the year of the last election and ending one year before the next election. In Table \@ref(tab:ratio-example), the Average Change value of $43.43$ equals the average Yearly Change over the years 2004 through 2007.

By operationalising service cuts in this manner, I aim at capturing the electorate’s retrospective vote, as voters judge the incumbent parties on past performances (see p. \pageref{passage:retrospective-vote}). When casting a ballot, I assume that they take into account the development of service quality since the last election. Service cuts lead to a decrease in service quality, which in turn can lead to a decrease in incumbent support.

```{r ratio-example, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.cap="ratio example"}

df <- ro_schools %>%
  dplyr::filter(NUTS_ID == "RO111") %>%
  dplyr::filter(year >=2004 & year <= 2008) %>%
  dplyr::select(year, schools, population, ratio_schools, ratio_schools_diff, average_ratio_schools_election_year) %>%
  mutate(ratio_schools = round(ratio_schools, 3)) %>%
  mutate(ratio_schools_diff = round(ratio_schools_diff, 3)) %>%
  mutate(average_ratio_schools_election_year = round(average_ratio_schools_election_year, 3)) %>%

  mutate(change_sign = case_when(ratio_schools_diff > 0 ~ "+",
                                 T ~ "")) %>%
  mutate(ratio_schools_diff = paste0(change_sign, ratio_schools_diff)) %>%
  mutate(ratio_schools_diff = gsub("-", "–", ratio_schools_diff)) %>%

  mutate(change_sign = case_when(average_ratio_schools_election_year > 0 ~ "+",
                                 T ~ "")) %>%

  mutate(average_ratio_schools_election_year = paste0(change_sign, average_ratio_schools_election_year)) %>%
  mutate(average_ratio_schools_election_year = gsub("-", "–", average_ratio_schools_election_year)) %>%

  mutate(average_ratio_schools_election_year = as.character(average_ratio_schools_election_year)) %>%
  mutate(average_ratio_schools_election_year = case_when(average_ratio_schools_election_year == "NA" ~ "",
                                                         T ~ average_ratio_schools_election_year)) %>%
  
  mutate(population = scales::number(population, big.mark = " ", decimal.mark = ".", accuracy = 1)) %>% # Add a space as the thousands separator
  dplyr::select(-change_sign) %>% 
  arrange(year) %>%
  dplyr::rename(Year = year,
                `Number of Schools` = schools,
                `Primary School Aged Population` = population,
                `Children per School` = ratio_schools,
                `Yearly Change in Children per School` = ratio_schools_diff,
                `Average Change in Children per School at E1` = average_ratio_schools_election_year)

kable(df,
      caption = "Operationalising service cuts to primary schools in Bihor County, Romania (NUTS3 ID: RO111)",
      format = "latex", booktabs = TRUE, escape = FALSE) %>%
  kable_styling(font_size = 9) %>%
  column_spec(column = 1, width = "0.8cm") %>%
  column_spec(column = c(2,4), width = "1.8cm") %>%
  column_spec(column = 3, width = "2cm") %>%
  column_spec(column = 5, width = "3cm") %>%
  column_spec(column = 6, width = "3.2cm") %>%
  add_footnote("Romanian national parliamentary elections took place in 2004 and in 2008. Source: Romanian National Institute of Statistics", notation = "none")

```

As described in the Theory section on p. \pageref{passage:theory-service-cuts}, the three service cuts under examination are school and hospital closures, as well as shutting down third places. I operationalise these service cuts as the average changing ratio of primary school aged children per primary school, the average changing ratio of people per hospital and the average changing ratio of people per third place, all within the same NUTS region.

#### Control Variables
In order to isolate the effect of service cuts on incumbent support, I control for external emigration, electoral volatility and economic performance within a NUTS region.

#### Emigration Rate
\label{passage:emigration-rate} When operationalising emigration rate, I base my approach on @dancygierEmigrationRadicalRight2022. I only take the number of departures from a NUTS region into account, in order to isolate the effect of emigration from other demographic factors that may influence the association between population decline and incumbent support [@dancygierEmigrationRadicalRight2022 5]. Also, I only take *external* emigration into account, i.e., migration abroad, and discard internal emigration between subnational regions. I focus on this because I aim to capture the effect of FoM on an EU member state.

I operationalise emigration rate as the sum of emigrants in a NUTS region [@dancygierEmigrationRadicalRight2022 17], divided by the average population in the same NUTS region between two consecutive national parliamentary elections. I then multiply this value by 1000 in order to retrieve the value per 1000 inhabitants. Table \@ref(tab:emigration-rate-example) displays this operationalisation with an example. 

This operationalisation aims at capturing a region’s entire population outflow and thus the impact of emigration on its remaining inhabitants. I use the sum of all emigrants from a region between two elections to capture the extent of emigration that a voter may perceive. Averaging the population is a means to find a middle value within a time period, in this case between two elections. Finally, I divide by the population and multiply by 1000 in order to calculate a region’s rate, thus allowing for comparisons between regions.

```{r emigration-rate-example, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.cap="emigration rate example"}

df <- ro %>% 
  dplyr::filter(NUTS_ID == "RO111") %>% 
  dplyr::filter(year >=2004 & year <= 2008) %>%
  dplyr::select(year, emigration, population, emigration_yearly_per_1000, emigration_election_year_per_1000) %>% 
  mutate(emigration_yearly_per_1000 = round(emigration_yearly_per_1000, 2)) %>% 
  mutate(emigration_election_year_per_1000 = round(emigration_election_year_per_1000, 2)) %>% 
  arrange(year) %>% 
  mutate(emigration_election_year_per_1000 = as.character(emigration_election_year_per_1000)) %>% 
  mutate(emigration_election_year_per_1000 = case_when(is.na(emigration_election_year_per_1000) == T ~ "",
                                                         T ~ emigration_election_year_per_1000)) %>% 
  mutate(population = scales::number(population, big.mark = " ", decimal.mark = ".", accuracy = 1)) %>% # Add a space as the thousands separator
  dplyr::rename(Year = year,
                `Number of Emigrants` = emigration,
                `Population Size` = population,
                `Yearly Emigration Rate` = emigration_yearly_per_1000,
                `Emigration Rate at E1` = emigration_election_year_per_1000)

kable(df,
      caption = "Operationalising emigration rate in Bihor County, Romania (NUTS3 ID: RO111)",
      format = "latex", booktabs = TRUE, escape = FALSE) %>% 
  kable_styling(font_size = 9) %>% 
  column_spec(column = 1, width = "0.8cm") %>% 
  column_spec(column = c(2,3), width = "1.8cm") %>% 
  # column_spec(column = c(5), width = "3cm") %>% 
  add_footnote("Romanian national parliamentary elections took place in 2004 and in 2008. Source: Romanian National Institute of Statistics", notation = "none")

```
  
#### Electoral Volatility
\label{passage:electoral-volatility} I operationalise electoral volatility by calculating the Pedersen Index [@pedersenElectoralVolatilityWestern1979] for each NUTS region in every election that data are available for. The Pedersen Index provides a value between 0 and 100 representing the electoral volatility from one election to the next. A value of 0 means there is no change in party vote share between elections. A value of 100 means the opposite, that all parties from the previous election lose all vote share in the next election and are replaced with new parties. Using data from @schraffEuropeanNUTSlevelElection2023, I calculate the Pedersen Index using the politicsR package [@silvaSilvadenissonPoliticsRV02023] in the R programming language. Table \@ref(tab:ro-pedersen-example) shows the electoral volatility between the 2008 and 2012 national parliamentary elections in an exemplary Romanian NUTS3 region.

In the context of calculating the Pedersen Index, it is important to point out how @schraffEuropeanNUTSlevelElection2023 define parties. Each party receives a unique ID based on @doringPartyFactsDatabase2019. However, defining a party as a single entity over time has its caveats [@doringPartyFactsDatabase2019 101]. E.g., how should two merging parties, party dissolutions or party renaming be dealt with? Within the scope of this thesis, I deem the most sensible option to follow the party coding provided by @schraffEuropeanNUTSlevelElection2023, who code parties in the following manner. Two separate parties have their own unique IDs, $P_1$ and $P_2$, at election $E_0$. If these two parties then merge to compete together in the next election $E_1$, this will result in a third, new party ID, $P_3$. This new ID is completely detached from the previous two IDs, thus no connection can be made between the IDs $P_1$, $P_2$ and $P_3$. A consequence of this approach is that all three IDs exist for both elections, $E_0$ and $E_1$. However, $P_1$ and $P_2$ have a vote share of zero percent in $E_1$. This is because $P_1$ and $P_2$ did not compete in $E_1$, as they competed together as $P_3$.

A real life example can be observed in Table \@ref(tab:ro-pedersen-example). In Bihor County, Romania, the National Liberal Party (Party ID: 481) reached $23.83\%$ of the vote in the 2008 election, the Alliance PSD+PC (Party ID: 6153) got $17.72\%$ of the vote. In 2011, these two parties merged to form the Social Liberal Union^[Source: https://en.wikipedia.org/wiki/National_Liberal_Party_(Romania)#Transition_from_USL_to_ACL_and_third_governing_experiences_(2010%E2%80%932020), retrieved 27 January 2024] (Party ID: 5941) and competed under this name in the 2012 election. Due to this merger, the National Liberal Party and the Alliance PSD+PC both display an election result of zero percent in the 2012 election. On the other hand, the Social Liberal Union has zero percent vote share in the 2008 election and $49.25\%$ vote share in the 2012 election.

Treating party mergers this way works if voters consider a new alliance of existing parties to be a novel party. This is however a very charitable assumption and most likely will not hold. In sum, this setup will result in a Pedersen Index that is too high, as parties that merge do not truly lose their entire electoral share from one election to the next. Developing a method to more precisely capture party landscape changes over time would be called for, but this is not feasible within the scope of this thesis. Thus I revert to using the unique party IDs provided by @schraffEuropeanNUTSlevelElection2023.

```{r ro-pedersen-example, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.cap="ro pedersen example"}
ro_volatility_2012 <- ro_volatility_table %>% 
  dplyr::select(-y_2004, -y_2016) %>% 
  group_by(nuts2016) %>% 
  dplyr::filter(!is.na(y_2008) | !is.na(y_2012)) %>% 
  mutate(y_2008 = case_when(is.na(y_2008) == T ~ 0,
                            T ~ y_2008)) %>% 
  mutate(y_2012 = case_when(is.na(y_2012) == T ~ 0,
                            T ~ y_2012)) %>% 
  mutate(diff = abs(y_2012 - y_2008)) %>% 
  mutate(volatility_2012 = volatility(y_2008, y_2012)) %>% 
  dplyr::filter(nuts2016 == "RO111") %>% 
  ungroup(nuts2016) %>% 
  dplyr::select(partyfacts_id, y_2008, y_2012, diff) %>% 
  mutate(y_2008 = round(y_2008, 3)) %>% 
  mutate(y_2012 = round(y_2012, 3)) %>% 
  mutate(diff = round(diff, 3)) %>% 
  arrange(-y_2008) %>% 
  mutate(partyfacts_id = as.character(partyfacts_id)) %>% 
  mutate(partyfacts_id = case_when(is.na(partyfacts_id) == T ~ "(Party ID missing)",
                                   T ~ partyfacts_id)) %>%
  dplyr::rename(`Party ID` = partyfacts_id,
                `Absolute Difference between E0 and E1` = diff,
                `Vote Share 2008 (E0) in %` = y_2008,
                `Vote Share 2012 (E1) in %` = y_2012)

kable(ro_volatility_2012,
      caption = "Calculating electoral volatility using the Pedersen Index in Bihor County, Romania (NUTS3 ID: RO111)",
      format = "latex", booktabs = TRUE, linesep = "") %>% 
  kable_styling(font_size = 9) %>% 
  column_spec(column = c(2,3), width = "3cm") %>%
  column_spec(column = 4, width = "3cm") %>%
  add_footnote("The Pedersen Index equals the sum of the absolute differences divided by 2 (Pedersen 1979). The sum of the absolute differences in Bihor County is 147, thus the Pedersen Index equals 73.5 for the year 2012. Party IDs for parties with very low vote shares are missing (Schraff et al. 2023, 572). Full party names can be accessed at partyfacts.herokuapp.com/data/partycodes/Party ID/. Source: Schraff et al. 2023", notation="none")
```

#### Economic Performance
I operationalise a region’s economic performance using two variables. The first variable is a region’s yearly gross domestic product (GDP) at current market prices in Euros per inhabitant. These data are provided by Eurostat.^[https://ec.europa.eu/eurostat/databrowser/view/nama_10r_3gdp__custom_9030433/default/table?lang=en (retrieved 16 December 2023)] I use the GDP per capita as a general indicator for a region’s economy under the assumption that this value captures the overall economic performance in a way that is comparable across different subnational regions.

The second variable I use to operationalise a region’s economic performance is its unemployment rate, provided by the Romanian Statistical Agency.^[Variable *SOM103A*: Unemployment rate by gender, macroregions, development regions and counties (http://statistici.insse.ro:8077/tempo-online/#/pages/tables/insse-table, retrieved 19 January 2024)] The unemployment rate is a common indicator to judge the incumbent party’s economic performance by, widely recognised and easily understood by voters.

In sum, both economic performance variables aim at capturing the overall economic mood that the electorate perceives and judges incumbent parties by when casting their vote.
