### Research Design
In this thesis I explicitly differentiate between internal emigration, i.e., emigration within a country, and external emigration, i.e., emigration from one country to another. I assume that the EU’s FoM makes emigration between member states easier. Based on this assumption, I focus only on external emigration in order to best isolate the impact of FoM on voting behaviour. I will measure the effects of service cuts on the lowest subnational NUTS level that data are available for. Emigration rates vary in different subnational regions in Poland and Romania (see Figure \@ref(fig:pl-ro-emigration-rate)). Taking these regions into account will account for this variation, creating more accurate models.

```{r pl-ro-emigration-rate, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Yearly Emigration Rate per 1000 Population"}

df <- cee_crude_emigration %>%
  dplyr::filter(country == c("RO", "PL")) %>% 
  dplyr::filter(year < 2020) %>% 
  mutate(country = case_when(country == "PL" ~ "Polish NUTS2 Regions",
                             country == "RO" ~ "Romanian NUTS3 Regions"))

ggplot(df, aes(x = year, y = emigration_yearly_per_1000, group = NUTS_ID)) +
  geom_line(color = "black") +
  geom_point(shape = 21, fill = "white", color = "black", size = 1) +
  facet_wrap(~ country,
             scales = "free_x",
             ncol = 2) +
  ylab(" ") +
  xlab(" ") +
  theme_minimal() +
  theme(
    text = element_text(family = "serif"),
        strip.text = element_text(hjust = 0) # Left aligns the facet titles
  )
```

### Overview of available data, how it was collected and transformed

<!-- Ideal Data Situation -->
In order to measure the effect of a country’s EU accession on emigration, ideal emigration data would be longitudinal, beginning several years before a country’s accession and would continue up until 2019, presumably after which the COVID-19 pandemic would interfere with emigration rates. Furthermore, the data would be available for the smallest subnational region (NUTS3), would explicitly differentiate between internal and external emigration and would be provided in the form of a rate per 1000 population.

Analogue to the emigration data, election results would ideally be available for each party within a NUTS3 region and for elections in the same time period that emigration data is available for. Furthermore, the data would ideally explicitly state if a party is in government, supports government or is in opposition. This makes it possible to differentiate between pro- and anti-incumbent support within a region.

However, the available data is not this extensives, as displayed in Table X (TO DO: data availability table).

```
Data Availability Table (example data)

Country | Emigration Data (Source) | Election Results (Source)    | NUTS              | … ???
PL      | 1998-2019  | 2004, 2008, …        | NUTS2: 17 regions
RO      | 1998-2019  | 2004, 2008, …        | NUTS3: 42 regions

```

<!-- Incumbent Party Data -->
The V-Party Dataset [@staffan.i.lindbergVPartyDatasetV22022] provides information on which parties are in or support the government.^[Variable “v2pagovsup”: Does this party support the government formed immediately after this election? (Codebook p. 24)]. I manually add further elections and parties to the data^[See https://github.com/fabianaiolfi/MA_Thesis/blob/main/scripts/01_anti_incumbent_vote/02_1_manual_adjustment.R for details] in order to match it to the emigration data time span and to minimise the number of missing values. Data on whether a party is incumbent or is based on the documented electoral history on their Wikipedia page.^[E.g., https://en.wikipedia.org/wiki/Greater_Romania_Party#Electoral_history (retrieved 15 January 2024)] This manual extension results in a more complete picture of anti-incumbent voting. However, I only manually add missing values of parties with a vote share equal to or larger than 4\% and a vote change between elections equal to or larger than 1\%. I do this in order to keep the scope of this thesis manageable.

In a next step, I adjust the V-Party Dataset and categorise parties that show government support leading up to an election as incumbent, even if they are outside of government. I do this because I assume a citizen who wants to cast an anti-incumbent vote will vote against any party that supports or sypmathises with incumbent parties.

<!-- Election Results Data -->
Subnational election results are provided by the European NUTS-level election dataset [@schraffEuropeanNUTSlevelElection2023]. I remove EU parliament elections from the data, leaving only national parliament election results. I also remove votes cast from abroad, parties named “Other” and elections before the year 1994 and after 2020. I calculate a party’s vote share by dividing the number of received votes by the total number of votes cast in the region, then multiply this by 100 to receive the regional vote share in percent. In a next step, I calculate a party’s regional vote change between two consecutive elections.

Using unique party IDs, I can combine the V-Party Dataset with the European NUTS-level election dataset. This combined dataset shows the regional vote change of incumbent parties. A random sample from this dataset is displayed in Table \@ref(tab:ned-vdem-sample).

<!-- Election Years -->
The years in which national parliamentary elections were held can be retrieved from the European NUTS-level election dataset. I manually add election years in order for the dataset to align with the emigration data. These missing years are retrieved from Wikipedia.^[E.g., https://en.wikipedia.org/wiki/2019_Polish_parliamentary_election or https://en.wikipedia.org/wiki/2016_Romanian_parliamentary_election (both retrieved 15 January 2024)]

```{r ned-vdem-sample, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.cap="Show snippet of n_v_dem_cee"}
df <- ned_v_dem_cee %>%
  drop_na(prev_incumbent) %>% 
  mutate(vote_share = round(vote_share, 2)) %>% 
  mutate(vote_change = round(vote_change, 2)) %>% 
  mutate(vote_change_sign = case_when(vote_change > 0 ~ "+",
                                      T ~ "")) %>%
  dplyr::select(year, country, regionname, nuts2016, unique_party_id, vote_share, vote_change_sign, vote_change, prev_incumbent) %>%
  mutate(vote_change = paste0(vote_change_sign, vote_change)) %>% 
  mutate(vote_change = gsub("-", "–", vote_change)) %>% 
  dplyr::rename(Year = year,
                Country = country,
                `Region Name` = regionname,
                `NUTS ID` = nuts2016,
                `Party ID` = unique_party_id,
                `Vote Share %` = vote_share,
                `Vote Change` = vote_change,
                `Incumbent?` = prev_incumbent,
                ) %>% 
  dplyr::select(-vote_change_sign) %>% 
  sample_n(5) %>% 
  arrange(Year)

kable(df, caption = "Random Sample of NED V-DEM CEE", format = "latex", booktabs = TRUE) %>% 
  kable_styling(font_size = 9) %>% 
  add_footnote("Full party name can be accessed through partyfacts.herokuapp.com/data/partycodes/Party ID/", notation="alphabet")
```

<!-- External emigration for all CEE -->
Data on external emigration are provided by national statistical offices. I define external emigration as the sum of people moving abroad from a NUTS region in a single year. I pick this absolute value because it carries more validity when it comes to measuring the impact of emigration than, e.g., net migration, a popular metric when measuring migration. Net migration has the disadvantage of including *immigration* in its calculation, thus hiding the number of people emigrating from a region.

<!-- Service Cuts -->
Data on service cuts at regional levels are provided by national statistical offices.

<!-- Service Cuts in Poland -->
The Polish government agency “Statistics Poland”^[https://bdl.stat.gov.pl/bdl/start, retrieved 20 November 2023] provides data on the number of primary schools at NUTS2 level^[Variable P1781: Primary schools for children, youth, and adults, by governing authority (https://bdl.stat.gov.pl/bdl/dane/podgrup/tablica, retrieved 20 November 2023)] and the size of the primary school population^[Variable P1341: Population by singular age and sex (https://bdl.stat.gov.pl/bdl/dane/podgrup/tablica, retrieved 20 November 2023)].

“Statistics Poland” also provides NUTS2 level data on the number of hospitals ^[Variable P2452: General hospitals (https://bdl.stat.gov.pl/bdl/dane/podgrup/tablica. retrieved 24 November 2023)] as well as the number of third places. ^[Variable P2505: Number of catering estabilishments [sic] (https://bdl.stat.gov.pl/bdl/dane/podgrup/tablica, retrieved 27 November 2023)]

<!-- Service Cuts in Romania -->
The Romanian government agency “National Institute of Statistics”^[https://insse.ro/cms/en, retrieved 15 January 2024] hosts data on the number of primary and lower-secondary schools^[Variable SCL101A: Education units, by categories of units, ownerships, macroregions, development regions and counties (http://statistici.insse.ro:8077/tempo-online/#/pages/tables/insse-table, retrieved 15 December 2023)] and the size of the population enrolled in these schools^[Variable SCL103E: Enrolled population, by level of education, urban/ rural area, macroregions, development regions and counties (http://statistici.insse.ro:8077/tempo-online/#/pages/tables/insse-table, retrieved 15 December 2023)]. Furthermore, the Institute also counts the number of hospitals^[Variable SCL101A: Sanitary units, by category of units, ownerships, macroregions, development regions and counties (http://statistici.insse.ro:8077/tempo-online/#/pages/tables/insse-table, retrieved 18 December 2023)] and third places.^[Variable INT101U: Active local units, by activity of national economy at level of CANE Rev.2 classes, size classes of number of employees, macroregions, development regions and counties (http://statistici.insse.ro:8077/tempo-online/#/pages/tables/insse-table, retrieved 15 December 2023)]

<!-- Romanian Classrooms as a proxy -->
As Figure X shows (TO DO), there is an inexplicable large drop in the number of Romanian schools in the early 2000s. I assume that along with schools closings, the Romanian “National Institute of Statistics” changed the way they counted single school entities. To counter this problem, I use the number of classrooms in primary and lower-secondary schools^[Variable SCL105C: Classrooms (school cabinets/amphiteaters), by level of education, macroregions, development regions and counties (http://statistici.insse.ro:8077/tempo-online/#/pages/tables/insse-table, retrieved 8 January 2024)] as a proxy for the number of school buildings. Figure X (TO DO) displays the number of classrooms, whereby a decrease is visible across most regions, but no sudden drop occurs. Thus I assume the number of classrooms to be an adequate proxy for the number of Romanian school buildings.
