### Preliminary Analysis

- in order to test the hypotheses, three assumptions have to hold
- one, does a country’s accession to the EU affect external emigration?
- two, does external emigration affect service cuts?
- three, does external emigration affect incumbent vote share?
- i examine these three assumptions by setting up two-way fixed effects linear regression models using the data described above
- I apply Two-way fixed effects models as this helps model data that varies across both regions and years
- the Polish model use NUTS2 and Romanian models NUTS3 regions as well as years as fixed effects, allowing me to remove any potentially confounding unobserved region- or time-specific factors.

```{r, child = 'preliminary_analysis_accession_emigration.Rmd'}
```

```{r, child = 'preliminary_analysis_emigration_service_cuts.Rmd'}
```

<!-- ```{r, child = 'xxx.Rmd'} -->
<!-- # Does emigration cause grievances? -->
<!-- ``` -->

```{r, child = 'preliminary_analysis_emigration_incumbent_vote_share.Rmd'}
```
