#### Does EU Accession Affect Emigration?  

```{r echo=FALSE, message=FALSE, warning=FALSE}
pl_emigration <- pl %>% 
  group_by(year) %>% 
  summarise(total_emigration = sum(emigration),
            total_population = sum(population)) %>% 
  mutate(emigration_rate = total_emigration / total_population * 1000) %>% 
  dplyr::filter(year <= 2020)

ro_emigration <- ro %>% 
  group_by(year) %>% 
  summarise(total_emigration = sum(emigration),
            total_population = sum(population)) %>% 
  mutate(emigration_rate = total_emigration / total_population * 1000) %>% 
  dplyr::filter(year <= 2020)

pl_emigration_plot <- ggplot(pl_emigration, aes(x = year, y = emigration_rate)) +
  geom_vline(xintercept = 2004, linetype = "dashed") +
  geom_bar(stat = "identity") +
  xlab("") + ylab("") +
  labs(subtitle = "Poland") +
  theme_minimal() +
  theme(text = element_text(family = "serif")) +
  # Remove minor y-axis grid lines
  theme(panel.grid.minor.y = element_blank()) +
  # Remove vertical grid lines
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        plot.caption = element_text(hjust = 0, vjust = 0, size = 8)) +
  labs(caption = "Dashed line: Year of EU accession. Sources: Statistics Poland, Romanian National Institute of Statistics")

ro_emigration_plot <- ggplot(ro_emigration, aes(x = year, y = emigration_rate)) +
  geom_vline(xintercept = 2007, linetype = "dashed") +
  geom_bar(stat = "identity") +
  xlab("") + ylab("") +
  labs(subtitle = "Romania") +
  theme_minimal() +
  theme(text = element_text(family = "serif")) +
  # Remove minor y-axis grid lines
  theme(panel.grid.minor.y = element_blank()) +
  # Remove 0.25 and 0.75 y-axis values
  scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2, 2.5, 3)) +
  # Remove vertical grid lines
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        plot.caption = element_text(hjust = 0, vjust = 0, size = 8)) +
  labs(caption = " ")

# Chow test
chow_ro <- sctest(ro_emigration$emigration_rate ~ ro_emigration$year,
       type = "Chow",
       point = which(ro_emigration$year == 2007))

chow_pl <- sctest(pl_emigration$emigration_rate ~ pl_emigration$year,
       type = "Chow",
       point = which(pl_emigration$year == 2004))
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Function to perform Chow test for a single region
perform_chow_test <- function(data, breakpoint_year) {
  sctest(emigration_yearly_per_1000 ~ year, data = data,
         type = "Chow",
         point = which(data$year == breakpoint_year))
}

# Applying the function to each NUTS2 region in Poland
results <- pl %>%
  group_by(NUTS_ID) %>%
  do(chow_test = perform_chow_test(., 2004))

# Plot results
p_values <- sapply(results$chow_test, function(x) x$p.value)
p_values <- as.data.frame(p_values)

p_values_significant <- p_values %>% dplyr::filter(p_values <= 0.05)

p_values_pl <- ggplot(p_values, aes(x = p_values)) +
  geom_boxplot() +
  labs(subtitle = "Poland") +
  xlab("p-value") +
  scale_y_continuous(breaks = NULL) + # Remove y-axis numbers
  xlim(0, 0.01) +
  theme_minimal() +
  theme(text = element_text(family = "serif"))

# Applying the function to each NUTS3 region in Romania
results <- ro %>%
  group_by(NUTS_ID) %>%
  do(chow_test = perform_chow_test(., 2007))

# Plot results
p_values <- sapply(results$chow_test, function(x) x$p.value)
p_values <- as.data.frame(p_values)

p_values_significant <- p_values %>% dplyr::filter(p_values <= 0.05)

p_values_ro <- ggplot(p_values, aes(x = p_values)) +
  geom_boxplot() +
  labs(subtitle = "Romania") +
  xlab("p-value") +
  scale_y_continuous(breaks = NULL) + # Remove y-axis numbers
  xlim(0, 0.05) +
  theme_minimal() +
  theme(text = element_text(family = "serif"))
```

Gaining EU membership grants member states access to FoM, facilitating migration between member states. Does emigration from CEE EU member states increase after accession? I examine this question in two steps. In a first step, I plot external emigration from Poland and Romania over time. If the assumption that access to FoM leads to an increase in emigration holds, we should observe higher emigration rates after a country’s accession. Figure \@ref(fig:accession-emigration-bar) displays each country’s emigration rate over time, with the vertical dashed line marking the year of accession to the EU. In both Poland and Romania we can observe a constant emigration rate until joining the EU. Once a country joins, emigration rates increase. However, the increase is lagged by two years in Poland and four years in Romania. The two countries also display different developments of emigration after this initial increase. Poland’s emigration rate decreases constantly after a peak is reached in 2006, with another small increase around 2013 and 2014. Romania displays a constant increase in emigration after 2010. These two varying developments hint at other factors influencing emigration rates alongside access to FoM.

```{r accession-emigration-bar, echo=FALSE, fig.height=2, message=FALSE, warning=FALSE, fig.cap="External emigration over time"}
grid.arrange(pl_emigration_plot, ro_emigration_plot, ncol = 2)
```

In a second step, I perform a Chow-Test [@chowTestsEqualitySets1960] for each NUTS region in Poland and Romania to examine if EU accession creates a break in the rate of emigration over time. A country’s year of accession is set as the break point. The Chow-Test returns a p-value, with which I can determine if the development of the emigration rate over time displays a significant difference before and after accession. In this specific case, I receive a p-value for every NUTS region in Poland and Romania.

In Poland, 15 of the 17 NUTS2 regions, or around $88\%$, return p-values below conventional statistical significance ($p < 0.05$), while in Romania 38 of the 42 NUTS3 regions, about $90\%$, do. Figure \@ref(fig:accession-emigration-box) displays the distribution of the returned p-values in a box plot. The plots show that almost all regions in Poland and Romania display a significant break in emigration rate around EU accession.

```{r accession-emigration-box, echo=FALSE, fig.height=2, message=FALSE, warning=FALSE, fig.cap="Distribution of p-values for Chow Test"}
# Define a layout matrix
layout_matrix <- rbind(c(1, NA, 2))

# Arrange the plots with the specified layout
grid.arrange(p_values_pl, p_values_ro, layout_matrix = layout_matrix, widths = c(1, 0.25, 1))
```

Based on the visual interpretation of the bar plots and results from the Chow Test, I assume an association to exist between a country’s accession to the EU and a change in emigration rate. I assume this association to be positive, whereby accession increases emigration. This is likely due to a country’s access to FoM through EU membership. However, as the diverging emigration rates between Poland and Romania after joining the EU show, accession is not the only factor that influences external emigration.
