## Overview of XXXX

I now focus on the effects of service cuts on anti-incumbent voting in Poland. I start by providing an overview of the variables I later use in my models. NUTS2 data is used, because Poland only provides data at this level. 

\pagebreak

### Population in Millions in Poland at NUTS2 Level
The population has decreased in a majority of Polish NUTS2 regions. The four regions that have seen an increase are XXXXX.

worms_explainer

```{r echo=FALSE, fig.height = 8, message=FALSE, warning=FALSE}
pl_avg_population_change <- pl %>% 
  arrange(NUTS_ID, year) %>%
  group_by(NUTS_ID) %>%
  mutate(population_change = population - dplyr::lag(population),
         # Calculate the year-on-year percent increase. Multiplying by 100 gives the percent change.
         percent_change = (population_change / dplyr::lag(population)) * 100
         ) %>%
  summarise(
    avg_population_change = mean(population_change, na.rm = TRUE), # Average change in actual numbers
    avg_percent_change = mean(percent_change, na.rm = TRUE) # Average percent change
  )

pl %>% 
  left_join(pl_avg_population_change, by = "NUTS_ID") %>% 
  mutate(population = population / 1000000) %>% 
  mutate(NUTS_ID = reorder(NUTS_ID, -avg_percent_change)) %>% 
  ggplot(aes(x = year, y = population)) +
  geom_smooth(aes(color = avg_percent_change), method = "loess", span = worm_span, se = FALSE) +
  scale_color_gradient(name = "Average of yearly change in population (in %) ") +
  scale_x_continuous(breaks = c(1995, 2022)) +
  pl_plots_components
```

\pagebreak

### Primary Age Population in 1000 in Poland at NUTS2 Level
The primary age population, i.e., children aged seven to twelve, has seen a very similar development across regions apart from XXXx (PL91). This segment of the population decreased steadily until around the year 2010–2012, after which an uptick is visible.

worms_explainer

```{r echo=FALSE, fig.height = 8, message=FALSE, warning=FALSE}
pl_schools_avg_population_change <- pl_schools %>% 
  arrange(NUTS_ID, year) %>% # Arrange data by region and then by year
  group_by(NUTS_ID) %>% # Group data by region
  mutate(population_change = population - dplyr::lag(population),
         percent_change = (population_change / dplyr::lag(population)) * 100) %>% # Calculate year-on-year increase
  summarise(avg_population_change = mean(population_change, na.rm = TRUE),
            avg_percent_change = mean(percent_change, na.rm = TRUE)) # Calculate average increase

pl_schools %>% 
  left_join(pl_schools_avg_population_change, by = "NUTS_ID") %>% 
  mutate(population = population / 1000) %>% 
  mutate(avg_population_change = avg_population_change / 1000) %>% 
  mutate(NUTS_ID = reorder(NUTS_ID, -avg_percent_change)) %>% 
  ggplot(aes(x = year, y = population)) +
  geom_smooth(aes(color = avg_percent_change), method = "loess", span = worm_span, se = FALSE) +
  scale_color_gradient(name = "Average of yearly change in population (in 1000s) ") +
	scale_x_continuous(breaks = c(1998, 2022)) +
  pl_plots_components
```

\pagebreak

### External Emigration per 1000 Population from Poland at NUTS2 Level

Most regions display two peaks in external emigration around the years 2006 and 2013/2014. The first peak may be due to Poland’s accession to the EU in the year 2004. The second peak may be due to Polish citizens gaining access to the EU’s Freedom of Movement for all EU member states (xxxxxx check this).

worms_explainer

```{r echo=FALSE, fig.height = 8, message=FALSE, warning=FALSE}
pl_avg_emigration_change <- pl %>% 
  arrange(NUTS_ID, year) %>% # Arrange data by region and then by year
  group_by(NUTS_ID) %>% # Group data by region
  mutate(emigration_change = emigration_yearly_per_1000 - dplyr::lag(emigration_yearly_per_1000),
         percent_change = (emigration_change / dplyr::lag(emigration_yearly_per_1000)) * 100) %>% 
  summarise(avg_emigration_change = mean(emigration_change, na.rm = TRUE),
            avg_percent_change = mean(percent_change, na.rm = TRUE)) # Calculate average increase

pl %>% 
  left_join(pl_avg_emigration_change, by = "NUTS_ID") %>% 
  mutate(NUTS_ID = reorder(NUTS_ID, -avg_percent_change)) %>% 
  ggplot(aes(x = year, y = emigration_yearly_per_1000)) +
  geom_smooth(aes(color = avg_percent_change), method = "loess", span = worm_span, se = FALSE) +
  scale_color_gradient(name = "Average of yearly change in emigration ") +
	scale_x_continuous(breaks = c(1995, 2022)) +
  pl_plots_components
```

\pagebreak

## Number of Institutions
I provide separate overviews of the number of schools, hospitals and third places in Polish NUTS2 regions over time.

### Number of Schools in Poland at NUTS2 Level
Apart from XXXX (PL91), all Polish NUTS2 regions see an overall decrease in the number of schools, with some regions displaying an uptick between the years 2015 and 2017. This aligns with my assumption that emigration leads to less schools.

worms_explainer

```{r echo=FALSE, fig.height = 8, message=FALSE, warning=FALSE}
pl_schools_avg_change <- pl_schools %>%
  arrange(NUTS_ID, year) %>% # Arrange data by region and then by year
  group_by(NUTS_ID) %>% # Group data by region
  mutate(school_change = schools - dplyr::lag(schools),
         percent_change = (school_change / dplyr::lag(schools)) * 100) %>% 
  summarise(avg_school_change = mean(school_change, na.rm = TRUE),
            avg_percent_change = mean(percent_change, na.rm = TRUE))

pl_schools %>% 
  left_join(pl_schools_avg_change, by = "NUTS_ID") %>%
  mutate(NUTS_ID = reorder(NUTS_ID, -avg_percent_change)) %>%
  ggplot(aes(x = year, y = schools)) +
  geom_smooth(aes(color = avg_percent_change), method = "loess", span = worm_span, se = FALSE) +
  scale_color_gradient(name = "Average of yearly change in number of schools ") +
	scale_x_continuous(breaks = c(1998, 2022)) +
  pl_plots_components
```

\pagebreak

### Number of Hospitals in Poland at NUTS2 Level
```{r echo=FALSE, fig.height = 8, message=FALSE, warning=FALSE}
pl_hospitals_avg_change <- pl_hospitals %>%
  arrange(NUTS_ID, year) %>% # Arrange data by region and then by year
  group_by(NUTS_ID) %>% # Group data by region
  mutate(hospital_change = hospitals - dplyr::lag(hospitals),
         percent_change = (hospital_change / dplyr::lag(hospitals)) * 100) %>% 
  summarise(avg_hospital_change = mean(hospital_change, na.rm = TRUE),
            avg_percent_change = mean(percent_change, na.rm = TRUE))

pl_hospitals %>% 
  left_join(pl_hospitals_avg_change, by = "NUTS_ID") %>%
  mutate(NUTS_ID = reorder(NUTS_ID, -avg_percent_change)) %>%
  ggplot(aes(x = year, y = hospitals)) +
  geom_smooth(aes(color = avg_percent_change), method = "loess", span = worm_span, se = FALSE) +
  scale_color_gradient(name = "average of yearly change in number of hospitals ") +
  pl_plots_components
```
\pagebreak

### Number of Third Places in Poland at NUTS2 Level
```{r echo=FALSE, fig.height = 8, message=FALSE, warning=FALSE}
pl_third_places_avg_change <- pl_third_places %>%
  arrange(NUTS_ID, year) %>% # Arrange data by region and then by year
  group_by(NUTS_ID) %>% # Group data by region
  mutate(third_places_change = third_places - dplyr::lag(third_places),
         percent_change = (third_places_change / dplyr::lag(third_places)) * 100) %>% 
  summarise(avg_third_places_change = mean(third_places_change, na.rm = TRUE),
            avg_percent_change = mean(percent_change, na.rm = TRUE))

pl_third_places %>% 
  left_join(pl_third_places_avg_change, by = "NUTS_ID") %>%
  mutate(NUTS_ID = reorder(NUTS_ID, -avg_percent_change)) %>%
  ggplot(aes(x = year, y = third_places)) +
  geom_smooth(aes(color = avg_percent_change), method = "loess", span = worm_span, se = FALSE) +
  scale_color_gradient(name = "Average of yearly change in number of third places ") +
  pl_plots_components +
  theme(legend.position = "bottom")
```


\pagebreak

## Ratios

Also contrary to my expectations, all services that I investigate generally see a decrease in their ratios. This means that over time, there are less people per institution in a NUTS2 region. E.g., less children in a school or less citizens per hospital.

\pagebreak

### Number of Primary School Children per School in Poland at NUTS2 Level
```{r echo=FALSE, fig.height = 8, message=FALSE, warning=FALSE}
pl_ratio_schools_avg_change <- pl_schools %>%
  arrange(NUTS_ID, year) %>% # Arrange data by region and then by year
  group_by(NUTS_ID) %>% # Group data by region
  mutate(ratio_school_change = ratio_schools - dplyr::lag(ratio_schools),
         percent_change = (ratio_school_change / dplyr::lag(ratio_schools)) * 100) %>% 
  summarise(avg_ratio_school_change = mean(ratio_school_change, na.rm = TRUE),
            avg_percent_change = mean(percent_change, na.rm = TRUE))

pl_schools %>% 
  left_join(pl_ratio_schools_avg_change, by = "NUTS_ID") %>%
  mutate(NUTS_ID = reorder(NUTS_ID, -avg_percent_change)) %>%
  ggplot(aes(x = year, y = ratio_schools)) +
  geom_smooth(aes(color = avg_percent_change), method = "loess", span = worm_span, se = FALSE) +
  scale_color_gradient(name = "Average of yearly change in children per school ") +
  pl_plots_components
```
\pagebreak

### Number of People per Hospital in Poland at NUTS2 Level
```{r echo=FALSE, fig.height = 8, message=FALSE, warning=FALSE}
pl_ratio_hospitals_avg_change <- pl_hospitals %>%
  arrange(NUTS_ID, year) %>% # Arrange data by region and then by year
  group_by(NUTS_ID) %>% # Group data by region
  mutate(ratio_hospital_change = ratio_hospitals_all_population - dplyr::lag(ratio_hospitals_all_population),
         percent_change = (ratio_hospital_change / dplyr::lag(ratio_hospitals_all_population)) * 100) %>% 
  summarise(avg_ratio_hospital_change = mean(ratio_hospital_change, na.rm = TRUE),
            avg_percent_change = mean(percent_change, na.rm = TRUE))

pl_hospitals %>% 
  left_join(pl_ratio_hospitals_avg_change, by = "NUTS_ID") %>%
  mutate(NUTS_ID = reorder(NUTS_ID, -avg_percent_change)) %>%
  ggplot(aes(x = year, y = ratio_hospitals_all_population)) +
  geom_smooth(aes(color = avg_percent_change), method = "loess", span = worm_span, se = FALSE) +
  scale_color_gradient(name = "Average of yearly change in people per hospital ") +
  pl_plots_components
```

\pagebreak

### Number of People per Third Place in Poland at NUTS2 Level
```{r echo=FALSE, fig.height = 8, message=FALSE, warning=FALSE}
pl_ratio_third_places_avg_change <- pl_third_places %>%
  arrange(NUTS_ID, year) %>% # Arrange data by region and then by year
  group_by(NUTS_ID) %>% # Group data by region
  mutate(ratio_third_places_change = ratio_third_places - dplyr::lag(ratio_third_places),
         percent_change = (ratio_third_places_change / dplyr::lag(ratio_third_places)) * 100) %>% 
  summarise(avg_ratio_third_places_change = mean(ratio_third_places_change, na.rm = TRUE),
            avg_percent_change = mean(percent_change, na.rm = TRUE))

pl_third_places %>% 
  left_join(pl_ratio_third_places_avg_change, by = "NUTS_ID") %>%
  mutate(NUTS_ID = reorder(NUTS_ID, -avg_percent_change)) %>%
  ggplot(aes(x = year, y = ratio_third_places)) +
  geom_smooth(aes(color = avg_percent_change), method = "loess", span = worm_span, se = FALSE) +
  scale_color_gradient(name = "Average of yearly change in people per third place ") +
  pl_plots_components +
  theme(legend.position = "bottom")
```

\pagebreak

## Control Variables

Both remittances and GDP show an increase over time. I use these data as control variables in my models. Data on remittances are only available at the national level.

### Remittance Inflows (US$ million) for all of Poland
```{r echo=FALSE, fig.height=3, message=FALSE, warning=FALSE}
ggplot(pl_remittances, aes(x = year, y = remittances)) +
  geom_line(color = "light grey") +
  geom_smooth(aes(group = 1), method = "loess", color = "red", se = FALSE, size = 0.5) + 
  scale_y_continuous(labels = label_comma()) + ylab("") + xlab("") + theme_minimal()
```

### GDP at Current Market Prices in Poland at NUTS2 Level (Euros per Inhabitant)
```{r echo=FALSE, fig.height=3, message=FALSE, warning=FALSE}
ggplot(pl_gdp, aes(x = year, y = gdp, line = NUTS_ID)) +
  geom_line(color = "light grey") +
  geom_smooth(aes(group = 1), method = "loess", color = "red", se = FALSE, size = 0.5) + 
  scale_y_continuous(labels = label_comma()) + ylab("") + xlab("") + theme_minimal()
```
