## Modelling Anti-Incumbent Voting in Poland at NUTS2 Level using Ratios

```{r echo=FALSE, message=FALSE, warning=FALSE}
options(scipen = 0) # Enable scientific notation

anti_incumbent_vote <- ned_v_dem_cee %>% 
  dplyr::filter(country == "Poland") %>%
  dplyr::filter(prev_incumbent == T) %>%
  left_join(dplyr::select(pl, NUTS_ID, year, emigration_election_year_per_1000), by = c("year" = "year", "nuts2016" = "NUTS_ID"))

anti_incumbent_vote <- anti_incumbent_vote %>% 
  left_join(dplyr::select(pl_schools, -population), by = c("year" = "year", "nuts2016" = "NUTS_ID")) %>%
  left_join(dplyr::select(pl_hospitals, -population), by = c("year" = "year", "nuts2016" = "NUTS_ID")) %>% 
  left_join(dplyr::select(pl_third_places, -population), by = c("year" = "year", "nuts2016" = "NUTS_ID")) %>% 
  left_join(pl_remittances, by = "year") %>% 
  left_join(pl_gdp, by = c("year" = "year", "nuts2016" = "NUTS_ID"))
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Using ratio in election year
fe_lm_1 <- feols(vote_change ~
                   ratio_schools +
                   emigration_election_year_per_1000 +
                   remittances +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_2 <- feols(vote_change ~
                   ratio_hospitals_all_population +
                   emigration_election_year_per_1000 +
                   remittances +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_3 <- feols(vote_change ~
                   ratio_third_places +
                   emigration_election_year_per_1000 +
                   remittances +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_4 <- feols(vote_change ~
                   ratio_schools +
                   ratio_hospitals_all_population +
                   emigration_election_year_per_1000 +
                   remittances +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_5 <- feols(vote_change ~
                   ratio_schools +
                   ratio_third_places +
                   emigration_election_year_per_1000 +
                   remittances +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_6 <- feols(vote_change ~
                   ratio_hospitals_all_population +
                   ratio_third_places +
                   emigration_election_year_per_1000 +
                   remittances +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_7 <- feols(vote_change ~
                ratio_schools +
                ratio_hospitals_all_population +
                ratio_third_places +
                emigration_election_year_per_1000 +
                remittances +
                gdp | 
                nuts2016 + year,
              data = anti_incumbent_vote)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
modelsummary(list(fe_lm_1, fe_lm_2, fe_lm_3),
             title = "DV: Incumbent's Change in Vote Share in Poland",
             coef_map = c(ratio_schools = "Children per School",
                          ratio_hospitals_all_population = "People per Hospital",
                          ratio_third_places = "People per Third Place",
                          emigration_election_year_per_1000 = "Avg. Emigration between Elections",
                          gdp = "GDP"),
             stars = T,
             statistic = NULL)
```

I first model separate ratios on incumbent vote change while controlling for emigration, GDP and remittances. Models 1, 2 and 3 in Table 3 each use a ratio separately. NUTS2 regions and years are used as fixed effects in all models. The emigration rate is the rolling emigration average between two national elections per 1000 inhabitants. Remittances are removed due to collinearity.

Two coefficients point in the right direction but overall one variable is statistically significant at conventional levels. An increase in the number of children per school is associated with a decrease in the incumbent vote share. However, the p-value is high at conventional levels ($p = 0.197$). The ratio of people per hospital displays no effect on incumbent vote share change, while the number of people per third place has a very small yet statistically significant effect.

\pagebreak

```{r echo=FALSE, message=FALSE, warning=FALSE}
modelsummary(list(fe_lm_4, fe_lm_5, fe_lm_6, fe_lm_7),
             title = "DV: Incumbent's Change in Vote Share in Poland",
             coef_map = c(ratio_schools = "Children per School",
                          ratio_hospitals_all_population = "People per Hospital",
                          ratio_third_places = "People per Third Place",
                          emigration_election_year_per_1000 = "Avg. Emigration between Elections",
                          gdp = "GDP"),
             stars = T,
             statistic = NULL)
```

In a second step, I use different combinations of ratios (Table 4, models 1, 2 and 3) and finally combine all ratios in model 4 (see Table 4). Again, NUTS2 regions and years are used as fixed effects. Both school ratio and third places ratio display statistically significant and negative coefficients, which align with my theory. This means that an increase in the number of children per school and an increase of the number of people per third place is associated with a decrease in the incumbent vote share, while controlling for emigration and GDP. The third ratio, number of people per hospital, still shows no effect.

\pagebreak

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Using average ratio between elections
fe_lm_8 <- feols(vote_change ~
                   average_ratio_schools_election_year +
                   emigration_election_year_per_1000 +
                   remittances +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_9 <- feols(vote_change ~
                   average_ratio_hospitals_election_year +
                   emigration_election_year_per_1000 +
                   remittances +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_10 <- feols(vote_change ~
                   average_ratio_third_places_election_year +
                   emigration_election_year_per_1000 +
                   remittances +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_11 <- feols(vote_change ~
                   average_ratio_schools_election_year +
                   average_ratio_hospitals_election_year +
                   emigration_election_year_per_1000 +
                   remittances +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_12 <- feols(vote_change ~
                   average_ratio_schools_election_year +
                   average_ratio_third_places_election_year +
                   emigration_election_year_per_1000 +
                   remittances +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_13 <- feols(vote_change ~
                   average_ratio_hospitals_election_year +
                   average_ratio_third_places_election_year +
                   emigration_election_year_per_1000 +
                   remittances +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_14 <- feols(vote_change ~
                average_ratio_schools_election_year +
                average_ratio_hospitals_election_year +
                average_ratio_third_places_election_year +
                emigration_election_year_per_1000 +
                remittances +
                gdp | 
                nuts2016 + year,
              data = anti_incumbent_vote)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
modelsummary(list(fe_lm_8, fe_lm_9, fe_lm_10),
             title = "DV: Incumbent's Change in Vote Share in Poland",
             coef_map = c(average_ratio_schools_election_year = "Avg. Change of Children per School",
                          average_ratio_hospitals_election_year = "Avg. Change of People per Hospital",
                          average_ratio_third_places_election_year = "Avg. Change of People per Third Place",
                          emigration_election_year_per_1000 = "Avg. Emigration between Elections",
                          gdp = "GDP"),
             stars = T,
             statistic = NULL)
```

Using average ratio between elections…

\pagebreak

```{r echo=FALSE, message=FALSE, warning=FALSE}
modelsummary(list(fe_lm_11, fe_lm_12, fe_lm_13, fe_lm_14),
             title = "DV: Incumbent's Change in Vote Share in Poland",
             coef_map = c(average_ratio_schools_election_year = "Avg. Change of Children per School",
                          average_ratio_hospitals_election_year = "Avg. Change of People per Hospital",
                          average_ratio_third_places_election_year = "Avg. Change of People per Third Place",
                          emigration_election_year_per_1000 = "Avg. Emigration between Elections",
                          gdp = "GDP"),
             stars = T,
             statistic = NULL)
```

Using average ratio between elections…

\pagebreak

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Interactions
fe_lm_15 <- feols(vote_change ~
                   average_ratio_schools_election_year *
                   emigration_election_year_per_1000 +
                   remittances +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_16 <- feols(vote_change ~
                   average_ratio_hospitals_election_year *
                   emigration_election_year_per_1000 +
                   remittances +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_17 <- feols(vote_change ~
                   average_ratio_third_places_election_year *
                   emigration_election_year_per_1000 +
                   remittances +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
modelsummary(list(fe_lm_15, fe_lm_16, fe_lm_17),
             title = "DV: Incumbent's Change in Vote Share in Poland",
             coef_map = c(average_ratio_schools_election_year = "Avg. Change of Children per School",
                          `average_ratio_schools_election_year:emigration_election_year_per_1000` = "Avg. Change of Children per School × Avg. Emigration",
                          average_ratio_hospitals_election_year = "Avg. Change of People per Hospital",
                          `average_ratio_hospitals_election_year:emigration_election_year_per_1000` = "Avg. Change of People per Hospital × Avg. Emigration",
                          average_ratio_third_places_election_year = "Avg. Change of People per Third Place",
                          `average_ratio_third_places_election_year:emigration_election_year_per_1000` = "Avg. Change of People per Third Place × Avg. Emigration",
                          emigration_election_year_per_1000 = "Avg. Emigration between Elections",
                          gdp = "GDP"),
             stars = T,
             statistic = NULL)
```

Using interactions…
