## Modelling Anti-Incumbent Voting in Poland at NUTS2 Level using Ratios

```{r echo=FALSE, message=FALSE, warning=FALSE}
options(scipen = 0) # Enable scientific notation

# Calculate the average crude_net_migration for each NUTS_ID for every pair of consecutive years
average_emigration <- pl %>%
  arrange(NUTS_ID, year) %>%
  group_by(NUTS_ID) %>%
  # Take rolling average of past 2 years into account, ignore current year
  mutate(average_emigration = slider::slide_dbl(emigration_yearly_per_1000, mean, .before = 2, .after = -1, .complete = T)) %>%
  ungroup()

anti_incumbent_vote <- ned_v_dem_cee %>% 
  dplyr::filter(prev_incumbent == T) %>%
  left_join(average_emigration, by = c("year" = "year", "nuts2016" = "NUTS_ID")) %>% 
  drop_na(emigration_yearly_per_1000)

anti_incumbent_vote <- anti_incumbent_vote %>% 
  left_join(dplyr::select(pl_schools, -population), by = c("year" = "year", "nuts2016" = "NUTS_ID")) %>%
  left_join(dplyr::select(pl_hospitals, -population), by = c("year" = "year", "nuts2016" = "NUTS_ID")) %>% 
  left_join(dplyr::select(pl_third_places, -population), by = c("year" = "year", "nuts2016" = "NUTS_ID")) %>% 
  left_join(pl_remittances, by = "year") %>% 
  left_join(pl_gdp, by = c("year" = "year", "nuts2016" = "NUTS_ID")) %>% 
  dplyr::filter(nuts2016 != "PL91") # Remove potentially problematic NUTS2 region (Warszawski stołeczny)

fe_lm_1 <- feols(vote_change ~
                   ratio_schools +
                   emigration_election_year_per_1000 +
                   remittances +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_2 <- feols(vote_change ~
                   ratio_hospitals_all_population +
                   emigration_election_year_per_1000 +
                   remittances +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_3 <- feols(vote_change ~
                   ratio_third_places +
                   emigration_election_year_per_1000 +
                   remittances +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_4 <- feols(vote_change ~
                   ratio_schools +
                   ratio_hospitals_all_population +
                   emigration_election_year_per_1000 +
                   remittances +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_5 <- feols(vote_change ~
                   ratio_schools +
                   ratio_third_places +
                   emigration_election_year_per_1000 +
                   remittances +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_6 <- feols(vote_change ~
                   ratio_hospitals_all_population +
                   ratio_third_places +
                   emigration_election_year_per_1000 +
                   remittances +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_7 <- feols(vote_change ~
                ratio_schools +
                ratio_hospitals_all_population +
                ratio_third_places +
                emigration_election_year_per_1000 +
                remittances +
                gdp | 
                nuts2016 + year,
              data = anti_incumbent_vote)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
modelsummary(list(fe_lm_1, fe_lm_2, fe_lm_3),
             title = "DV: Incumbent's Change in Vote Share in Poland",
             coef_map = c(ratio_schools = "Children per School",
                          ratio_hospitals_all_population = "People per Hospital",
                          ratio_third_places = "People per Third Place",
                          emigration_election_year_per_1000 = "Avg. Emigration between Elections",
                          gdp = "GDP"),
             stars = T,
             statistic = NULL)
```

I first model separate ratios on incumbent vote change while controlling for emigration, GDP and remittances. Models 1, 2 and 3 in Table 3 each use a ratio separately. NUTS2 regions and years are used as fixed effects in all models. The emigration rate is the rolling emigration average between two national elections per 1000 inhabitants. Remittances are removed due to collinearity.

Some coefficients point in the right direction but overall no results are statistically significant at conventional levels. An increase in the number of children per school is associated with a decrease in the incumbent vote share. However, the p-value is high at conventional levels ($p = 0.197$). Both hospital and third places ratios have coefficients of 0, displaying no effect on incumbent vote share change.

\pagebreak

```{r echo=FALSE, message=FALSE, warning=FALSE}
modelsummary(list(fe_lm_4, fe_lm_5, fe_lm_6, fe_lm_7),
             title = "DV: Incumbent's Change in Vote Share in Poland",
             coef_map = c(ratio_schools = "Children per School",
                          ratio_hospitals_all_population = "People per Hospital",
                          ratio_third_places = "People per Third Place",
                          emigration_election_year_per_1000 = "Avg. Emigration between Elections",
                          gdp = "GDP"),
             stars = T,
             statistic = NULL)
```

In a second step, I use different combinations of ratios (Table 4, models 1, 2 and 3) and finally combine all ratios in model 4 (see Table 4). Again, NUTS2 regions and years are used as fixed effects. Both school ratio and third places ratio display statistically significant and negative coefficients, which align with my theory. This means that an increase in the number of children per school and an increase of the number of people per third place is associated with a decrease in the incumbent vote share, while controlling for emigration and GDP. The third ratio, number of people per hospital, still shows no effect.
