## Modelling Anti-Incumbent Voting in Poland at NUTS2 Level using Number of Institutions

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Models using number of institutions
fe_lm_1 <- feols(vote_change ~
                   schools +
                   emigration_election_year_per_1000 +
                   remittances +
                   volatility +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_2 <- feols(vote_change ~
                   hospitals +
                   emigration_election_year_per_1000 +
                   remittances +
                   volatility +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_3 <- feols(vote_change ~
                   third_places +
                   emigration_election_year_per_1000 +
                   remittances +
                   volatility +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_4 <- feols(vote_change ~
                schools +
                hospitals +
                third_places +
                emigration_election_year_per_1000 +
                remittances +
                volatility +
                gdp | 
                nuts2016 + year,
              data = anti_incumbent_vote)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
modelsummary(list(fe_lm_1, fe_lm_2, fe_lm_3, fe_lm_4),
             title = "DV: Incumbent's Change in Vote Share in Poland",
             coef_map = c(schools = "Number of Schools",
                          hospitals = "Number of Hospitals",
                          third_places = "Number of Third Places",
                          emigration_election_year_per_1000 = "Avg. Emigration between Elections",
                          volatility = "Volatility",
                          gdp = "GDP"),
             stars = T,
             statistic = NULL)
```

As an alternative to using number of people per institution as the independent variable, I instead use the actual number of institutions. I do this to examine if the closing of an institution has an effect on the incumbent’s vote share, regardless of the number of people who then use the institution.

In contrast to using ratios as independent variables, using actual numbers should result in positive coefficients in order to align with my theory. This is because an increase in the number of, e.g., schools, should be associated with an increase in the incumbent’s change in vote share – and reversely – a decrease in schools should be associated with a decrease in vote share.

A single result is statistically significant (Model 3), but its sign does not comply with my theory. The coefficient implies a decrease in incumbent vote share if the number of third places increases.

\pagebreak

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Using average ratio between elections
fe_lm_8 <- feols(vote_change ~
                   average_schools_diff_election_year +
                   emigration_election_year_per_1000 +
                   remittances +
                   volatility +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_9 <- feols(vote_change ~
                   average_hospitals_diff_election_year +
                   emigration_election_year_per_1000 +
                   remittances +
                   volatility +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_10 <- feols(vote_change ~
                   average_third_places_diff_election_year +
                   emigration_election_year_per_1000 +
                   remittances +
                   volatility +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_11 <- feols(vote_change ~
                average_schools_diff_election_year +
                average_hospitals_diff_election_year +
                average_third_places_diff_election_year +
                emigration_election_year_per_1000 +
                remittances +
                volatility +
                gdp | 
                nuts2016 + year,
              data = anti_incumbent_vote)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
modelsummary(list(fe_lm_8, fe_lm_9, fe_lm_10, fe_lm_11),
             title = "DV: Incumbent's Change in Vote Share in Poland",
             coef_map = c(average_schools_diff_election_year = "Avg. Change of Schools",
                          average_hospitals_diff_election_year = "Avg. Change of Hospitals",
                          average_third_places_diff_election_year = "Avg. Change of Third Places",
                          emigration_election_year_per_1000 = "Avg. Emigration between Elections",
                          volatility = "Volatility",
                          gdp = "GDP"),
             stars = T,
             statistic = NULL)
```

Analogue to Table 2, I calculate the average change of institutions between two parliamentary elections and use this as the independent variable. As in Table 4, I expect the coefficients to be positive. However, the only statistically significant coefficients are the average change of number of hospitals, and these are both negative. The negative coefficient implies a decrease in incumbent vote share when the number of hospitals increases since the last election.

\pagebreak

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Interactions
fe_lm_15 <- feols(vote_change ~
                   average_schools_diff_election_year *
                   emigration_election_year_per_1000 +
                   remittances +
                   volatility +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_16 <- feols(vote_change ~
                   average_hospitals_diff_election_year *
                   emigration_election_year_per_1000 +
                   remittances +
                   volatility +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_17 <- feols(vote_change ~
                   average_third_places_diff_election_year *
                   emigration_election_year_per_1000 +
                   remittances +
                   volatility +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
modelsummary(list(fe_lm_15, fe_lm_16, fe_lm_17),
             title = "DV: Incumbent's Change in Vote Share in Poland",
             coef_map = c(#average_schools_diff_election_year = "Avg. Change of Schools",
                          `average_schools_diff_election_year:emigration_election_year_per_1000` = "Avg. Change of Schools × Avg. Emigration",
                          #average_hospitals_diff_election_year = "Avg. Change Hospitals",
                          `average_hospitals_diff_election_year:emigration_election_year_per_1000` = "Avg. Change Hospitals × Avg. Emigration",
                          #average_third_places_diff_election_year = "Avg. Change of Third Places",
                          `average_third_places_diff_election_year:emigration_election_year_per_1000` = "Avg. Change of Third Places × Avg. Emigration",
                          emigration_election_year_per_1000 = "Avg. Emigration between Elections",
                          volatility = "Volatility",
                          gdp = "GDP"),
             stars = T,
             statistic = NULL)
```

Analogue to Table 3, I add interaction terms between average change in number of institutions and average emigration. No results are statistically significant.
