## Modelling Anti-Incumbent Voting in Poland at NUTS2 Level using Number of Institutions

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Models using number of institutions
options(scipen = 0) # Enable scientific notation

# Calculate the average crude_net_migration for each NUTS_ID for every pair of consecutive years
average_emigration <- pl %>%
  arrange(NUTS_ID, year) %>%
  group_by(NUTS_ID) %>%
  # Take rolling average of past 2 years into account, ignore current year
  mutate(average_emigration = slider::slide_dbl(emigration_yearly_per_1000, mean, .before = 2, .after = -1, .complete = T)) %>%
  ungroup()

anti_incumbent_vote <- ned_v_dem_cee %>% 
  dplyr::filter(prev_incumbent == T) %>%
  left_join(average_emigration, by = c("year" = "year", "nuts2016" = "NUTS_ID")) %>% 
  drop_na(emigration_yearly_per_1000)

anti_incumbent_vote <- anti_incumbent_vote %>% 
  left_join(dplyr::select(pl_schools, -population), by = c("year" = "year", "nuts2016" = "NUTS_ID")) %>%
  left_join(dplyr::select(pl_hospitals, -population), by = c("year" = "year", "nuts2016" = "NUTS_ID")) %>% 
  left_join(dplyr::select(pl_third_places, -population), by = c("year" = "year", "nuts2016" = "NUTS_ID")) %>% 
  left_join(pl_remittances, by = "year") %>% 
  left_join(pl_gdp, by = c("year" = "year", "nuts2016" = "NUTS_ID"))

fe_lm_1 <- feols(vote_change ~
                   schools +
                   emigration_election_year_per_1000 +
                   remittances +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_2 <- feols(vote_change ~
                   hospitals +
                   emigration_election_year_per_1000 +
                   remittances +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_3 <- feols(vote_change ~
                   third_places +
                   emigration_election_year_per_1000 +
                   remittances +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_4 <- feols(vote_change ~
                   schools +
                   hospitals +
                   emigration_election_year_per_1000 +
                   remittances +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_5 <- feols(vote_change ~
                   schools +
                   third_places +
                   emigration_election_year_per_1000 +
                   remittances +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_6 <- feols(vote_change ~
                   hospitals +
                   third_places +
                   emigration_election_year_per_1000 +
                   remittances +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_7 <- feols(vote_change ~
                schools +
                hospitals +
                third_places +
                emigration_election_year_per_1000 +
                remittances +
                gdp | 
                nuts2016 + year,
              data = anti_incumbent_vote)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
modelsummary(list(fe_lm_1, fe_lm_2, fe_lm_3),
             title = "DV: Incumbent's Change in Vote Share in Poland",
             coef_map = c(schools = "Number of Schools",
                          hospitals = "Number of Hospitals",
                          third_places = "Number of Third Places",
                          emigration_election_year_per_1000 = "Avg. Emigration between Elections",
                          gdp = "GDP"),
             stars = T,
             statistic = NULL)

modelsummary(list(fe_lm_4, fe_lm_5, fe_lm_6, fe_lm_7),
             title = "DV: Incumbent's Change in Vote Share in Poland",
             coef_map = c(schools = "Number of Schools",
                          hospitals = "Number of Hospitals",
                          third_places = "Number of Third Places",
                          emigration_election_year_per_1000 = "Avg. Emigration between Elections",
                          gdp = "GDP"),
             stars = T,
             statistic = NULL)
```

As an alternative to using number of people per inhabitants as the independent variable, I instead use the actual number of institutions. I do this to examine if the closing of an institution has an effect on the incumbent’s vote share, regardless of the number of people who then use the institution.

Analog to Tables 3 and 4, I first model each institution seperately (Table 5) and then combine the different independent variables (Table 6). In constrast to using ratios as independent variables, using actual numbers should result in positive coefficients in order to align with my theory. This is because an increase in the number of, e.g., schools should be associated with an increase in the incumbent’s change in vote share – and reversely – a decrease in schools is associated with a decrease in vote share.

In sum, no model shows statisticlly significant results and the coefficients signs mostly do not align with my theory.

\pagebreak

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
modelsummary(list(fe_lm_4, fe_lm_5, fe_lm_6, fe_lm_7),
             title = "DV: Incumbent's Change in Vote Share in Poland",
             coef_map = c(schools = "Number of Schools",
                          hospitals = "Number of Hospitals",
                          third_places = "Number of Third Places",
                          emigration_election_year_per_1000 = "Avg. Emigration between Elections",
                          gdp = "GDP"),
             stars = T,
             statistic = NULL)
```
