## Population and Emigration

### Population in Millions in Romania at NUTS3 Level

`r worms_explainer`

```{r echo=FALSE, fig.height = 8, message=FALSE, warning=FALSE}
ro_avg_population_change <- ro_population %>% 
  arrange(NUTS_ID, year) %>%
  group_by(NUTS_ID) %>%
  mutate(population_change = population - dplyr::lag(population),
         # Calculate the year-on-year percent increase. Multiplying by 100 gives the percent change.
         percent_change = (population_change / dplyr::lag(population)) * 100
         ) %>%
  summarise(
    avg_population_change = mean(population_change, na.rm = TRUE), # Average change in actual numbers
    avg_percent_change = mean(percent_change, na.rm = TRUE) # Average percent change
  )

ro_population %>% 
  left_join(ro_avg_population_change, by = "NUTS_ID") %>% 
  mutate(population = population / 1000000) %>% 
  mutate(NUTS_ID = reorder(NUTS_ID, -avg_percent_change)) %>% 
  ggplot(aes(x = year, y = population)) +
  geom_smooth(aes(color = avg_percent_change), method = "loess", span = worm_span, se = FALSE) +
  scale_color_gradient(name = "Average of yearly change in population (in %) ") +
  scale_x_continuous(breaks = c(2000, 2023)) +
  ro_plots_components
```

\pagebreak

### Primary Age Population in 1000s in Romania at NUTS3 Level

`r worms_explainer`

```{r echo=FALSE, fig.height = 8, message=FALSE, warning=FALSE}
ro_schools_avg_population_change <- ro_schools %>% 
  arrange(NUTS_ID, year) %>% # Arrange data by region and then by year
  group_by(NUTS_ID) %>% # Group data by region
  mutate(population_change = population - dplyr::lag(population),
         percent_change = (population_change / dplyr::lag(population)) * 100) %>% # Calculate year-on-year increase
  summarise(avg_population_change = mean(population_change, na.rm = TRUE),
            avg_percent_change = mean(percent_change, na.rm = TRUE)) # Calculate average increase

ro_schools %>% 
  left_join(ro_schools_avg_population_change, by = "NUTS_ID") %>% 
  mutate(population = population / 1000) %>% 
  mutate(NUTS_ID = reorder(NUTS_ID, -avg_percent_change)) %>% 
  ggplot(aes(x = year, y = population)) +
  geom_smooth(aes(color = avg_percent_change), method = "loess", span = worm_span, se = FALSE) +
  scale_color_gradient(name = "Average of yearly change in population (in 1000s) ") +
  scale_x_continuous(breaks = c(1995, 2022)) +
  ro_plots_components
```

\pagebreak

### Emigration per 1000 Inhabitants from Romania at NUTS3 Level

`r worms_explainer`

```{r echo=FALSE, fig.height = 8, message=FALSE, warning=FALSE}
ro_avg_emigration_change <- ro %>% 
  arrange(NUTS_ID, year) %>% # Arrange data by region and then by year
  group_by(NUTS_ID) %>% # Group data by region
  mutate(emigration_change = emigration_yearly_per_1000 - dplyr::lag(emigration_yearly_per_1000),
         percent_change = (emigration_change / dplyr::lag(emigration_yearly_per_1000)) * 100) %>% 
  summarise(avg_emigration_change = mean(emigration_change, na.rm = TRUE),
            avg_percent_change = mean(percent_change, na.rm = TRUE)) # Calculate average increase

ro %>% 
  left_join(ro_avg_emigration_change, by = "NUTS_ID") %>% 
  mutate(NUTS_ID = reorder(NUTS_ID, -avg_percent_change)) %>% 
  ggplot(aes(x = year, y = emigration_yearly_per_1000)) +
  geom_smooth(aes(color = avg_percent_change), method = "loess", span = worm_span, se = FALSE) +
  scale_color_gradient(name = "Average of yearly change in emigration ") +
  scale_x_continuous(breaks = c(2000, 2022)) +
  ro_plots_components
```

\pagebreak

## Number of Institutions

### Number of Schools in Romania at NUTS3 Level

Many regions see a sudden drop in number of schools during the beginning of the 2000s, the reason for which is currently unclear.

`r worms_explainer`

```{r echo=FALSE, fig.height = 8, message=FALSE, warning=FALSE}
ro_schools_avg_change <- ro_schools %>%
  arrange(NUTS_ID, year) %>% # Arrange data by region and then by year
  group_by(NUTS_ID) %>% # Group data by region
  mutate(school_change = schools - dplyr::lag(schools),
         percent_change = (school_change / dplyr::lag(schools)) * 100) %>% 
  summarise(avg_school_change = mean(school_change, na.rm = TRUE),
            avg_percent_change = mean(percent_change, na.rm = TRUE))

ro_schools %>% 
  left_join(ro_schools_avg_change, by = "NUTS_ID") %>%
  mutate(NUTS_ID = reorder(NUTS_ID, -avg_percent_change)) %>%
  ggplot(aes(x = year, y = schools)) +
  geom_smooth(aes(color = avg_percent_change), method = "loess", span = worm_span, se = FALSE) +
  scale_color_gradient(name = "Average of yearly change in number of schools ") +
  scale_x_continuous(breaks = c(1990, 2022)) +
  ro_plots_components
```

\pagebreak

### Number of Classrooms in Romania at NUTS3 Level

`r worms_explainer`

```{r echo=FALSE, fig.height = 8, message=FALSE, warning=FALSE}
ro_classrooms_avg_change <- ro_schools %>%
  arrange(NUTS_ID, year) %>% # Arrange data by region and then by year
  group_by(NUTS_ID) %>% # Group data by region
  mutate(classroom_change = classrooms - dplyr::lag(classrooms),
         percent_change = (classroom_change / dplyr::lag(classrooms)) * 100) %>% 
  summarise(avg_classroom_change = mean(classroom_change, na.rm = TRUE),
            avg_percent_change = mean(percent_change, na.rm = TRUE))

ro_schools %>% 
  left_join(ro_classrooms_avg_change, by = "NUTS_ID") %>%
  mutate(NUTS_ID = reorder(NUTS_ID, -avg_percent_change)) %>%
  ggplot(aes(x = year, y = classrooms)) +
  geom_smooth(aes(color = avg_percent_change), method = "loess", span = worm_span, se = FALSE) +
  scale_color_gradient(name = "Average of yearly change in number of classrooms ") +
  scale_x_continuous(breaks = c(1990, 2022)) +
  ro_plots_components
```

\pagebreak

### Number of Hospitals in Romania at NUTS3 Level

`r worms_explainer`

```{r echo=FALSE, fig.height = 8, message=FALSE, warning=FALSE}
ro_hospitals_avg_change <- ro_hospitals %>%
  arrange(NUTS_ID, year) %>% # Arrange data by region and then by year
  group_by(NUTS_ID) %>% # Group data by region
  mutate(hospital_change = hospitals - dplyr::lag(hospitals),
         percent_change = (hospital_change / dplyr::lag(hospitals)) * 100) %>% 
  summarise(avg_hospital_change = mean(hospital_change, na.rm = TRUE),
            avg_percent_change = mean(percent_change, na.rm = TRUE))

ro_hospitals %>% 
  left_join(ro_hospitals_avg_change, by = "NUTS_ID") %>%
  mutate(NUTS_ID = reorder(NUTS_ID, -avg_percent_change)) %>%
  ggplot(aes(x = year, y = hospitals)) +
  geom_smooth(aes(color = avg_percent_change), method = "loess", span = worm_span, se = FALSE) +
  scale_color_gradient(name = "average of yearly change in number of hospitals ") +
  scale_x_continuous(breaks = c(1990, 2022)) +
  ro_plots_components
```
\pagebreak

### Number of Third Places in Romania at NUTS3 Level

`r worms_explainer`

```{r echo=FALSE, fig.height = 8, message=FALSE, warning=FALSE}
ro_third_places_avg_change <- ro_third_places %>%
  arrange(NUTS_ID, year) %>% # Arrange data by region and then by year
  group_by(NUTS_ID) %>% # Group data by region
  mutate(third_places_change = third_places - dplyr::lag(third_places),
         percent_change = (third_places_change / dplyr::lag(third_places)) * 100) %>% 
  summarise(avg_third_places_change = mean(third_places_change, na.rm = TRUE),
            avg_percent_change = mean(percent_change, na.rm = TRUE))

ro_third_places %>% 
  left_join(ro_third_places_avg_change, by = "NUTS_ID") %>%
  mutate(NUTS_ID = reorder(NUTS_ID, -avg_percent_change)) %>%
  ggplot(aes(x = year, y = third_places)) +
  geom_smooth(aes(color = avg_percent_change), method = "loess", span = worm_span, se = FALSE) +
  scale_color_gradient(name = "Average of yearly change in number of third places ") +
  scale_x_continuous(breaks = c(2008, 2021)) +
  ro_plots_components
```

\pagebreak

## Ratios of People per Institution

In a next step, I look at the number of people per institution within a region. For example, I count the number of primary school children per school on average, or the number of a region’s inhabitants per hospital.

The following plots do not show a consistent trends across all regions in Romania. A possible next step could be to categorise NUTS3 regions, e.g., into urban and rural regions, or regions that border neighbouring countries, in an attempt to find groups with similar trends.

\pagebreak

### Number of Primary School Children per School in Romania at NUTS3 Level

We can generally observe an increase in children per school in Romania. This increase seems to follow the sudden decrease in schools at the beginning of the 2000s. 

`r worms_explainer`

```{r echo=FALSE, fig.height = 8, message=FALSE, warning=FALSE}
ro_ratio_schools_avg_change <- ro_schools %>%
  arrange(NUTS_ID, year) %>% # Arrange data by region and then by year
  group_by(NUTS_ID) %>% # Group data by region
  mutate(ratio_school_change = ratio_schools - dplyr::lag(ratio_schools),
         percent_change = (ratio_school_change / dplyr::lag(ratio_schools)) * 100) %>% 
  summarise(avg_ratio_school_change = mean(ratio_school_change, na.rm = TRUE),
            avg_percent_change = mean(percent_change, na.rm = TRUE))

ro_schools %>% 
  left_join(ro_ratio_schools_avg_change, by = "NUTS_ID") %>%
  mutate(NUTS_ID = reorder(NUTS_ID, -avg_percent_change)) %>%
  ggplot(aes(x = year, y = ratio_schools)) +
  geom_smooth(aes(color = avg_percent_change), method = "loess", span = worm_span, se = FALSE) +
  scale_color_gradient(name = "Average of yearly change in children per school ") +
  scale_x_continuous(breaks = c(1995, 2022)) +
  ro_plots_components
```

\pagebreak

### Number of Primary School Children per School in Romania at NUTS3 Level (Map)

```{r echo=FALSE, fig.height = 8, message=FALSE, warning=FALSE}
# Load the shapefile for maps
# https://ec.europa.eu/eurostat/web/gisco/geodata/reference-data/administrative-units-statistical-units/nuts
nuts3_shape_2016 <- st_read(here("data", "NUTS_RG_10M_2016_3035_shp", "NUTS_RG_10M_2016_3035.shp"))

# Merge the data
merged_data_2016 <- left_join(nuts3_shape_2016, ro_ratio_schools_avg_change, by = "NUTS_ID")
merged_data_2016_filter <- merged_data_2016 %>% dplyr::filter(LEVL_CODE == 3 & CNTR_CODE == "RO")
```

```{r echo=FALSE, fig.height = 8, message=FALSE, warning=FALSE}
ggplot() +
  geom_sf(data = merged_data_2016_filter, aes(fill = avg_percent_change), color = "#333333", lwd = 0.1) +
  scale_fill_viridis_c(guide = guide_colourbar(direction = "horizontal", barheight = 0.5, title.position = "bottom")) +
  theme_minimal() +
  theme(panel.grid = element_blank(), # remove grid
        axis.title = element_blank(), # remove axis labels
        axis.text = element_blank(),  # remove axis text
        legend.position = "bottom",   # place legend at bottom
        legend.text = element_text(size = 12),  # Increase the font size for legend items
        legend.title = element_text(size = 12, face = "bold")  # Increase the font size for the legend title
  ) +
  labs(fill = "Average Percent Change\nof Children per School")
```

\pagebreak

### Number of Primary School Children per Classroom in Romania at NUTS3 Level

`r worms_explainer`

```{r echo=FALSE, fig.height = 8, message=FALSE, warning=FALSE}
ro_ratio_classrooms_avg_change <- ro_schools %>%
  arrange(NUTS_ID, year) %>% # Arrange data by region and then by year
  group_by(NUTS_ID) %>% # Group data by region
  mutate(ratio_classroom_change = ratio_classrooms - dplyr::lag(ratio_classrooms),
         percent_change = (ratio_classroom_change / dplyr::lag(ratio_classrooms)) * 100) %>% 
  summarise(avg_ratio_classroom_change = mean(ratio_classroom_change, na.rm = TRUE),
            avg_percent_change = mean(percent_change, na.rm = TRUE))

ro_schools %>% 
  left_join(ro_ratio_classrooms_avg_change, by = "NUTS_ID") %>%
  mutate(NUTS_ID = reorder(NUTS_ID, -avg_percent_change)) %>%
  ggplot(aes(x = year, y = ratio_classrooms)) +
  geom_smooth(aes(color = avg_percent_change), method = "loess", span = worm_span, se = FALSE) +
  scale_color_gradient(name = "Average of yearly change in children per classroom ") +
  scale_x_continuous(breaks = c(1995, 2022)) +
  ro_plots_components
```

\pagebreak

### Number of People in 1000s per Hospital in Romania at NUTS3 Level

`r worms_explainer`

```{r echo=FALSE, fig.height = 8, message=FALSE, warning=FALSE}
ro_ratio_hospitals_avg_change <- ro_hospitals %>%
  arrange(NUTS_ID, year) %>% # Arrange data by region and then by year
  group_by(NUTS_ID) %>% # Group data by region
  mutate(ratio_hospital_change = ratio_hospitals - dplyr::lag(ratio_hospitals),
         percent_change = (ratio_hospital_change / dplyr::lag(ratio_hospitals)) * 100) %>% 
  summarise(avg_ratio_hospital_change = mean(ratio_hospital_change, na.rm = TRUE),
            avg_percent_change = mean(percent_change, na.rm = TRUE))

ro_hospitals %>% 
  mutate(ratio_hospitals = ratio_hospitals / 1000) %>% 
  left_join(ro_ratio_hospitals_avg_change, by = "NUTS_ID") %>%
  mutate(NUTS_ID = reorder(NUTS_ID, -avg_percent_change)) %>%
  ggplot(aes(x = year, y = ratio_hospitals)) +
  geom_smooth(aes(color = avg_percent_change), method = "loess", span = worm_span, se = FALSE) +
  scale_color_gradient(name = "Average of yearly change in people per hospital ") +
  scale_x_continuous(breaks = c(2000, 2022)) +
  ro_plots_components
```

\pagebreak

### Number of People in 1000s per Hospital in Romania at NUTS3 Level (Map)

```{r echo=FALSE, fig.height = 8, message=FALSE, warning=FALSE}
# Load the shapefile for maps
# https://ec.europa.eu/eurostat/web/gisco/geodata/reference-data/administrative-units-statistical-units/nuts
nuts3_shape_2016 <- st_read(here("data", "NUTS_RG_10M_2016_3035_shp", "NUTS_RG_10M_2016_3035.shp"))

# Merge the data
merged_data_2016 <- left_join(nuts3_shape_2016, ro_ratio_hospitals_avg_change, by = "NUTS_ID")
merged_data_2016_filter <- merged_data_2016 %>% dplyr::filter(LEVL_CODE == 3 & CNTR_CODE == "RO")
```

```{r echo=FALSE, fig.height = 8, message=FALSE, warning=FALSE}
ggplot() +
  geom_sf(data = merged_data_2016_filter, aes(fill = avg_percent_change), color = "#333333", lwd = 0.1) +
  scale_fill_viridis_c(guide = guide_colourbar(direction = "horizontal", barheight = 0.5, title.position = "bottom")) +
  theme_minimal() +
  theme(panel.grid = element_blank(), # remove grid
        axis.title = element_blank(), # remove axis labels
        axis.text = element_blank(),  # remove axis text
        legend.position = "bottom",   # place legend at bottom
        legend.text = element_text(size = 12),  # Increase the font size for legend items
        legend.title = element_text(size = 12, face = "bold")  # Increase the font size for the legend title
  ) +
  labs(fill = "Average Percent Change\nof People per Hospital")
```

\pagebreak

### Number of People in 1000s per Third Place in Romania at NUTS3 Level

`r worms_explainer`

```{r echo=FALSE, fig.height = 8, message=FALSE, warning=FALSE}
ro_ratio_third_places_avg_change <- ro_third_places %>%
  arrange(NUTS_ID, year) %>% # Arrange data by region and then by year
  group_by(NUTS_ID) %>% # Group data by region
  mutate(ratio_third_places_change = ratio_third_places - dplyr::lag(ratio_third_places),
         percent_change = (ratio_third_places_change / dplyr::lag(ratio_third_places)) * 100) %>% 
  summarise(avg_ratio_third_places_change = mean(ratio_third_places_change, na.rm = TRUE),
            avg_percent_change = mean(percent_change, na.rm = TRUE))

ro_third_places %>% 
  mutate(ratio_third_places = ratio_third_places / 1000) %>% 
  left_join(ro_ratio_third_places_avg_change, by = "NUTS_ID") %>%
  mutate(NUTS_ID = reorder(NUTS_ID, -avg_percent_change)) %>%
  ggplot(aes(x = year, y = ratio_third_places)) +
  geom_smooth(aes(color = avg_percent_change), method = "loess", span = worm_span, se = FALSE) +
  scale_color_gradient(name = "Average of yearly change in people per third place ") +
  scale_x_continuous(breaks = c(2008, 2020)) +
  ro_plots_components
```

\pagebreak

### Number of People in 1000s per Third Place in Romania at NUTS3 Level (Map)

```{r echo=FALSE, fig.height = 8, message=FALSE, warning=FALSE}
# Load the shapefile for maps
# https://ec.europa.eu/eurostat/web/gisco/geodata/reference-data/administrative-units-statistical-units/nuts
nuts3_shape_2016 <- st_read(here("data", "NUTS_RG_10M_2016_3035_shp", "NUTS_RG_10M_2016_3035.shp"))

# Merge the data
merged_data_2016 <- left_join(nuts3_shape_2016, ro_ratio_third_places_avg_change, by = "NUTS_ID")
merged_data_2016_filter <- merged_data_2016 %>% dplyr::filter(LEVL_CODE == 3 & CNTR_CODE == "RO")
```

```{r echo=FALSE, fig.height = 8, message=FALSE, warning=FALSE}
ggplot() +
  geom_sf(data = merged_data_2016_filter, aes(fill = avg_percent_change), color = "#333333", lwd = 0.1) +
  scale_fill_viridis_c(guide = guide_colourbar(direction = "horizontal", barheight = 0.5, title.position = "bottom")) +
  theme_minimal() +
  theme(panel.grid = element_blank(), # remove grid
        axis.title = element_blank(), # remove axis labels
        axis.text = element_blank(),  # remove axis text
        legend.position = "bottom",   # place legend at bottom
        legend.text = element_text(size = 12),  # Increase the font size for legend items
        legend.title = element_text(size = 12, face = "bold")  # Increase the font size for the legend title
  ) +
  labs(fill = "Average Percent Change\nof People per Third Place")
```

\pagebreak

## Control Variables

I use these data as control variables in my models. Data on remittances are only available at the national level.

### Remittance Inflows (US$ million) for all of Romania
```{r echo=FALSE, fig.height=3, message=FALSE, warning=FALSE}
ggplot(ro_remittances, aes(x = year, y = remittances)) +
  geom_line() +
  geom_point(shape = 21, fill = "white") +
  scale_y_continuous(labels = label_comma()) + ylab("") + xlab("") + theme_minimal()
```

### GDP at Current Market Prices in Romania at NUTS3 Level (Euros per Inhabitant)
Each line represents a NUTS3 region in Romania. Bucharest (RO321) displays a consistently higher GDP than all other regions.  

```{r echo=FALSE, fig.height=3, message=FALSE, warning=FALSE}
ggplot(ro_gdp, aes(x = year, y = gdp, line = NUTS_ID)) +
  geom_line() +
  geom_point(shape = 21, fill = "white") +
  scale_y_continuous(labels = label_comma()) + ylab("") + xlab("") + theme_minimal()
```
