## Modelling Anti-Incumbent Voting in Romania at NUTS3 Level using Ratios

```{r echo=FALSE, message=FALSE, warning=FALSE}
options(scipen = 0) # Enable scientific notation

anti_incumbent_vote <- ned_v_dem_cee %>% 
  dplyr::filter(country == "Romania") %>%
  dplyr::filter(prev_incumbent == T) %>%
  left_join(dplyr::select(ro, NUTS_ID, year, emigration_election_year_per_1000), by = c("year" = "year", "nuts2016" = "NUTS_ID"))

anti_incumbent_vote <- anti_incumbent_vote %>% 
  left_join(dplyr::select(ro_schools, -population), by = c("year" = "year", "nuts2016" = "NUTS_ID")) %>%
  left_join(dplyr::select(ro_hospitals, -population), by = c("year" = "year", "nuts2016" = "NUTS_ID")) %>% 
  left_join(dplyr::select(ro_third_places, -population), by = c("year" = "year", "nuts2016" = "NUTS_ID")) %>% 
  left_join(ro_remittances, by = "year") %>% 
  left_join(ro_gdp, by = c("year" = "year", "nuts2016" = "NUTS_ID")) %>% 
  left_join(ro_volatility, by = c("year", "nuts2016"))
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Using ratio in election year
fe_lm_1 <- feols(vote_change ~
                   ratio_schools +
                   emigration_election_year_per_1000 +
                   remittances +
                   volatility +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_2 <- feols(vote_change ~
                   ratio_hospitals +
                   emigration_election_year_per_1000 +
                   remittances +
                   volatility +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_3 <- feols(vote_change ~
                   ratio_third_places +
                   emigration_election_year_per_1000 +
                   remittances +
                   volatility +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_4 <- feols(vote_change ~
                ratio_schools +
                ratio_hospitals +
                ratio_third_places +
                emigration_election_year_per_1000 +
                remittances +
                volatility +
                gdp | 
                nuts2016 + year,
              data = anti_incumbent_vote)
```

```{r ro-fe-ratio, echo=FALSE, message=FALSE, warning=FALSE}
modelsummary(list(fe_lm_1, fe_lm_2, fe_lm_3, fe_lm_4),
             title = "DV: Incumbent's Change in Vote Share in Romania",
             coef_map = c(ratio_schools = "Children per School",
                          ratio_hospitals = "People per Hospital",
                          ratio_third_places = "People per Third Places",
                          emigration_election_year_per_1000 = "Avg. Emigration between Elections",
                          volatility = "Volatility",
                          gdp = "GDP"),
             stars = T,
             statistic = NULL)
```

I model separate ratios on incumbent vote change while controlling for emigration, volatility and GDP. Models 1, 2 and 3 in Table \@ref(tab:ro-fe-ratio) each use a ratio separately, Model 4 combines all ratios. NUTS3 regions and years are used as fixed effects in all models. The emigration rate is the rolling emigration average between two national parliamentary elections per 1000 inhabitants. No independent variable is statistically significant.

A highlight is the emigration control variable, which displays a relatively strong and statistically significant effect. Higher levels of emigration between elections is associated with a decrease in incumbent vote share across all models.

\pagebreak

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Interactions
fe_lm_a <- feols(vote_change ~
                   ratio_schools *
                   emigration_election_year_per_1000 +
                   remittances +
                   volatility +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_b <- feols(vote_change ~
                   ratio_hospitals *
                   emigration_election_year_per_1000 +
                   remittances +
                   volatility +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_c <- feols(vote_change ~
                   ratio_third_places *
                   emigration_election_year_per_1000 +
                   remittances +
                   volatility +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)
```

```{r ro-fe-ratio-interaction, echo=FALSE, message=FALSE, warning=FALSE}
modelsummary(list(fe_lm_a, fe_lm_b, fe_lm_c),
             title = "DV: Incumbent's Change in Vote Share in Romania",
             coef_map = c(#average_ratio_schools_election_year = "Avg. Change of Children per School",
                          `ratio_schools:emigration_election_year_per_1000` = "Children per School × Avg. Emigration",
                          #average_ratio_hospitals_election_year = "Avg. Change of People per Hospital",
                          `ratio_hospitals:emigration_election_year_per_1000` = "People per Hospital × Avg. Emigration",
                          #average_ratio_third_places_election_year = "Avg. Change of People per Third Place",
                          `ratio_third_places:emigration_election_year_per_1000` = "People per Third Place × Avg. Emigration",
                          emigration_election_year_per_1000 = "Avg. Emigration between Elections",
                          volatility = "Volatility",
                          gdp = "GDP"),
             stars = T,
             statistic = NULL)
```

I add a separate interaction term between every ratio and average emigration. No results are statistically significant.

\pagebreak

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Using average ratio between elections
fe_lm_8 <- feols(vote_change ~
                   average_ratio_schools_election_year +
                   emigration_election_year_per_1000 +
                   remittances +
                   volatility +
                   gdp |
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_8a <- feols(vote_change ~
                   average_ratio_classrooms_election_year +
                   emigration_election_year_per_1000 +
                   volatility +
                   gdp |
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_9 <- feols(vote_change ~
                   average_ratio_hospitals_election_year +
                   emigration_election_year_per_1000 +
                   remittances +
                   volatility +
                   gdp |
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_10 <- feols(vote_change ~
                   average_ratio_third_places_election_year +
                   emigration_election_year_per_1000 +
                   remittances +
                   volatility +
                   gdp |
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_11 <- feols(vote_change ~
                average_ratio_schools_election_year +
                average_ratio_hospitals_election_year +
                average_ratio_third_places_election_year +
                emigration_election_year_per_1000 +
                remittances +
                volatility +
                gdp |
                nuts2016 + year,
              data = anti_incumbent_vote)

fe_lm_11a <- feols(vote_change ~
                average_ratio_classrooms_election_year +
                average_ratio_hospitals_election_year +
                average_ratio_third_places_election_year +
                emigration_election_year_per_1000 +
                volatility +
                gdp |
                nuts2016 + year,
              data = anti_incumbent_vote)
```

```{r ro-fe-avg-ratio, echo=FALSE, message=FALSE, warning=FALSE}
modelsummary(list(fe_lm_8, fe_lm_8a, fe_lm_9, fe_lm_10, fe_lm_11, fe_lm_11a),
             title = "DV: Incumbent's Change in Vote Share in Romania",
             coef_map = c(average_ratio_schools_election_year = "Avg. Change of Children per School",
                          average_ratio_classrooms_election_year = "Avg. Change of Children per Classroom",
                          average_ratio_hospitals_election_year = "Avg. Change of People per Hospital",
                          average_ratio_third_places_election_year = "Avg. Change of People per Third Place",
                          emigration_election_year_per_1000 = "Avg. Emigration between Elections",
                          volatility = "Volatility",
                          gdp = "GDP"),
             stars = T,
             statistic = NULL)
```

Table \@ref(tab:ro-fe-avg-ratio) uses different independent variables compared to Table \@ref(tab:ro-fe-ratio). Instead of using the ratio in an election year, I calculate the average ratio _change_ between two national parliamentary elections. As in Table \@ref(tab:ro-fe-ratio), no independent variables are statistically significant.

\pagebreak

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Interactions
fe_lm_15 <- feols(vote_change ~
                   average_ratio_schools_election_year *
                   emigration_election_year_per_1000 +
                   remittances +
                   volatility +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_15a <- feols(vote_change ~
                   average_ratio_classrooms_election_year *
                   emigration_election_year_per_1000 +
                   remittances +
                   volatility +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_16 <- feols(vote_change ~
                   average_ratio_hospitals_election_year *
                   emigration_election_year_per_1000 +
                   remittances +
                   volatility +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_17 <- feols(vote_change ~
                   average_ratio_third_places_election_year *
                   emigration_election_year_per_1000 +
                   remittances +
                   volatility +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)
```

```{r ro-fe-avg-ratio-interaction, echo=FALSE, message=FALSE, warning=FALSE}
modelsummary(list(fe_lm_15, fe_lm_15a, fe_lm_16, fe_lm_17),
             title = "DV: Incumbent's Change in Vote Share in Romania",
             coef_map = c(#average_ratio_schools_election_year = "Avg. Change of Children per School",
                          `average_ratio_schools_election_year:emigration_election_year_per_1000` = "Avg. Change of Children per School × Avg. Emigration",
                          `average_ratio_classrooms_election_year:emigration_election_year_per_1000` = "Avg. Change of Children per Classroom × Avg. Emigration",
                          #average_ratio_hospitals_election_year = "Avg. Change of People per Hospital",
                          `average_ratio_hospitals_election_year:emigration_election_year_per_1000` = "Avg. Change of People per Hospital × Avg. Emigration",
                          #average_ratio_third_places_election_year = "Avg. Change of People per Third Place",
                          `average_ratio_third_places_election_year:emigration_election_year_per_1000` = "Avg. Change of People per Third Place × Avg. Emigration",
                          emigration_election_year_per_1000 = "Avg. Emigration between Elections",
                          volatility = "Volatility",
                          gdp = "GDP"),
             stars = T,
             statistic = NULL)
```

I add a separate interaction term between every average change in ratio and average emigration. Results show that the change in ratio of children per school and the change of people per hospital are statistically significant and have negative coefficients. This indicates that the negative impact of increased children per school and people per hospital on the incumbent’s vote share is moderated by the level of emigration. The average change of people per third place shows no effect. Also, as emigration increases, the negative effect of children per classroom on vote change becomes less negative.
