## Modelling Anti-Incumbent Voting in Romania at NUTS3 Level using Number of Institutions

```{r echo=FALSE, message=FALSE, warning=FALSE}
fe_lm_1 <- feols(vote_change ~
                   schools +
                   emigration_election_year_per_1000 +
                   remittances +
                   volatility +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_2 <- feols(vote_change ~
                   hospitals +
                   emigration_election_year_per_1000 +
                   remittances +
                   volatility +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_3 <- feols(vote_change ~
                   third_places +
                   emigration_election_year_per_1000 +
                   remittances +
                   volatility +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_4 <- feols(vote_change ~
                schools +
                hospitals +
                third_places +
                emigration_election_year_per_1000 +
                remittances +
                volatility +
                gdp | 
                nuts2016 + year,
              data = anti_incumbent_vote)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
modelsummary(list(fe_lm_1, fe_lm_2, fe_lm_3, fe_lm_4),
             title = "DV: Incumbent's Change in Vote Share in Romania",
             coef_map = c(schools = "Number of Schools",
                          hospitals = "Number of Hospitals",
                          third_places = "Number of Third Places",
                          emigration_election_year_per_1000 = "Avg. Emigration between Elections",
                          volatility = "Volatility",
                          gdp = "GDP"),
             stars = T,
             statistic = NULL)
```

As an alternative to using number of people per institution as the independent variable, I instead use the actual number of institutions. I do this to examine if the closing of an institution has an effect on the incumbent’s vote share, regardless of the number of people who then use the institution.

Analogue to Tables 9 and 10, I first model each institution separately (Table 11) and then combine the different independent variables (Table 12). In contrast to using ratios as independent variables, using actual numbers should result in positive coefficients in order to align with my theory. This is because an increase in the number of, e.g., schools, should be associated with an increase in the incumbent’s change in vote share – and reversely – a decrease in schools is associated with a decrease in vote share.

Using the number of schools as an independent variable is statistically significant across all models and always displays a positive coefficient (see Tables 11 and 12). This means that an decrease in the number of schools is associated with a decrease in the incumbent's vote share. In short, less schools in a region is bad for incumbent parties. All models also display relatively high R2 values, meaning that all models can explain a large proportion of the dependent variable's variance.

\pagebreak

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Using average diff between elections
fe_lm_8 <- feols(vote_change ~
                   average_schools_diff_election_year +
                   emigration_election_year_per_1000 +
                   remittances +
                   volatility +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_9 <- feols(vote_change ~
                   average_hospitals_diff_election_year +
                   emigration_election_year_per_1000 +
                   remittances +
                   volatility +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_10 <- feols(vote_change ~
                   average_third_places_diff_election_year +
                   emigration_election_year_per_1000 +
                   remittances +
                   volatility +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_11 <- feols(vote_change ~
                average_schools_diff_election_year +
                average_hospitals_diff_election_year +
                average_third_places_diff_election_year +
                emigration_election_year_per_1000 +
                remittances +
                volatility +
                gdp | 
                nuts2016 + year,
              data = anti_incumbent_vote)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
modelsummary(list(fe_lm_8, fe_lm_9, fe_lm_10, fe_lm_11),
             title = "DV: Incumbent's Change in Vote Share in Romania",
             coef_map = c(average_schools_diff_election_year = "Avg. Change of Schools",
                          average_hospitals_diff_election_year = "Avg. Change of Hospitals",
                          average_third_places_diff_election_year = "Avg. Change of Third Places",
                          emigration_election_year_per_1000 = "Avg. Emigration between Elections",
                          volatility = "Volatility",
                          gdp = "GDP"),
             stars = T,
             statistic = NULL)
```

Using average diff between elections…

\pagebreak

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Interactions
fe_lm_15 <- feols(vote_change ~
                   average_schools_diff_election_year *
                   emigration_election_year_per_1000 +
                   remittances +
                   volatility +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_16 <- feols(vote_change ~
                   average_hospitals_diff_election_year *
                   emigration_election_year_per_1000 +
                   remittances +
                   volatility +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)

fe_lm_17 <- feols(vote_change ~
                   average_third_places_diff_election_year *
                   emigration_election_year_per_1000 +
                   remittances +
                   volatility +
                   gdp | 
                   nuts2016 + year,
                 data = anti_incumbent_vote)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
modelsummary(list(fe_lm_15, fe_lm_16, fe_lm_17),
             title = "DV: Incumbent's Change in Vote Share in Romania",
             coef_map = c(#average_schools_diff_election_year = "Avg. Change of Schools",
                          `average_schools_diff_election_year:emigration_election_year_per_1000` = "Avg. Change of Schools × Avg. Emigration",
                          #average_hospitals_diff_election_year = "Avg. Change Hospitals",
                          `average_hospitals_diff_election_year:emigration_election_year_per_1000` = "Avg. Change Hospitals × Avg. Emigration",
                          #average_third_places_diff_election_year = "Avg. Change of Third Places",
                          `average_third_places_diff_election_year:emigration_election_year_per_1000` = "Avg. Change of Third Places × Avg. Emigration",
                          emigration_election_year_per_1000 = "Avg. Emigration between Elections",
                          volatility = "Volatility",
                          gdp = "GDP"),
             stars = T,
             statistic = NULL)
```

Using interactions…
