## Modelling with Truncated Data

Do models improve when truncating parts of the data? I truncate the data in three different ways to explore if results vary. Model 1 uses data where observations with a small vote change, between $-0.5$ and $0.5$, are removed. This is done to see if extreme vote change values lead to more significant results. Model 2 uses data in which observations with extremely high (over $30\%$) and extremely low (under $5\%$) vote shares are removed. Model 3 uses data where high levels of emigration rates (above 2 per 1000 inhabitants) are removed. Both models 2 and 3 attempt to remove outliers in order to build models with higher validity.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Truncate Vote Change
# Models using ratios
cutoff <- 0.5
anti_incumbent_vote_truncated <- anti_incumbent_vote %>% 
  dplyr::filter(vote_change > cutoff | vote_change < cutoff * -1)

lm_truncate_vote_change <- feols(vote_change ~
                                   ratio_schools +
                                   ratio_hospitals +
                                   ratio_third_places +
                                   emigration_election_year_per_1000 +
                                   remittances +
                                   gdp | 
                                   nuts2016 + year,
                                 data = anti_incumbent_vote_truncated)

# Truncate Vote Share
anti_incumbent_vote_truncated <- anti_incumbent_vote %>% 
  dplyr::filter(vote_share < 30 & vote_share > 5)

lm_truncate_vote_share <- feols(vote_change ~
                                  ratio_schools +
                                  ratio_hospitals +
                                  ratio_third_places +
                                  emigration_election_year_per_1000 +
                                  remittances +
                                  gdp | 
                                  nuts2016 + year,
                                data = anti_incumbent_vote_truncated)

# Truncate Emigration Rate
anti_incumbent_vote_truncated <- anti_incumbent_vote %>% 
  dplyr::filter(emigration_election_year_per_1000 < 2)

lm_truncate_emigration_rate <- feols(vote_change ~
                                       ratio_schools +
                                       ratio_hospitals +
                                       ratio_third_places +
                                       emigration_election_year_per_1000 +
                                       remittances +
                                       gdp | 
                                       nuts2016 + year,
                                     data = anti_incumbent_vote_truncated)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
modelsummary(list(lm_truncate_vote_change, lm_truncate_vote_share, lm_truncate_emigration_rate),
             title = "DV: Incumbent's Change in Vote Share in Romania (Truncated Data)",
             coef_map = c(ratio_schools = "Children per School",
                          ratio_hospitals = "People per Hospital",
                          ratio_third_places = "People per Third Places",
                          emigration_election_year_per_1000 = "Avg. Emigration between Elections",
                          gdp = "GDP"),
             stars = T,
             statistic = NULL)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Truncate Vote Change
# Models using number of institutions
cutoff <- 0.5
anti_incumbent_vote_truncated <- anti_incumbent_vote %>% 
  dplyr::filter(vote_change > cutoff | vote_change < cutoff * -1)

lm_truncate_vote_change <- feols(vote_change ~
                                   schools +
                                   hospitals +
                                   third_places +
                                   emigration_election_year_per_1000 +
                                   remittances +
                                   gdp | 
                                   nuts2016 + year,
                                 data = anti_incumbent_vote_truncated)

# Truncate Vote Share
anti_incumbent_vote_truncated <- anti_incumbent_vote %>% 
  dplyr::filter(vote_share < 30 & vote_share > 5)

lm_truncate_vote_share <- feols(vote_change ~
                                  schools +
                                  hospitals +
                                  third_places +
                                  emigration_election_year_per_1000 +
                                  remittances +
                                  gdp | 
                                  nuts2016 + year,
                                data = anti_incumbent_vote_truncated)

# Truncate Emigration Rate
anti_incumbent_vote_truncated <- anti_incumbent_vote %>% 
  dplyr::filter(emigration_election_year_per_1000 < 2)

lm_truncate_emigration_rate <- feols(vote_change ~
                                       schools +
                                       hospitals +
                                       third_places +
                                       emigration_election_year_per_1000 +
                                       remittances +
                                       gdp | 
                                       nuts2016 + year,
                                     data = anti_incumbent_vote_truncated)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
modelsummary(list(lm_truncate_vote_change, lm_truncate_vote_share, lm_truncate_emigration_rate),
             title = "DV: Incumbent's Change in Vote Share in Romania (Truncated Data)",
             coef_map = c(schools = "Number of Schools",
                          hospitals = "Number of Hospitals",
                          third_places = "Number of Third Places",
                          emigration_election_year_per_1000 = "Avg. Emigration between Elections",
                          gdp = "GDP"),
             stars = T,
             statistic = NULL)
```

Model results using truncated data are shows in Tables 13 and 14. The models in Table 13 use ratios as independent variables, models in Table 14 use numbers of institutions.

None of the truncated models show particularly promising results, apart model 1 when using actual numbers of institutions (see Table 14). As with the non-truncated data, the number of schools and incumbent change in vote share are positively associated, with a decrease in schools going hand in hand with a decrease in incumbent vote share. All other coefficients from all other models using truncated data do not display statistically significant results. Also as with the non-truncated data, the control variable for emigration displays highly significant and strong effects, with an increase in emigration being associated with a decrease in incubment vote share.
